### 2020-12-19 070-supplementary-materials.Rmd

# Supplementary Materials {.unnumbered #misc}

We provide an open-source implementation of our methodology as the `bigsimr` R package, hosted on github, https://schisslergroup.github.io/bigsimr/.

```{r, out.width='5%', fig.align='center'}
knitr::include_graphics('images/hex-bigsimr.png')
```

# Acknowledgment(s) {.unnumbered #acknowledgments}

The authors gratefully acknowledge Heather Knudson's graphic design for the `bigsimr` R Package.
The results published here are in whole or part based upon data generated by the TCGA Research Network: https://www.cancer.gov/tcga.

# Disclosure statement {.unnumbered #coi}

The authors report no conflict of interest.

# Funding {.unnumbered #funding}

Research reported in this publication was supported by grants from NV INBRE and the National Institutes of General Medical Sciences (NIH 5U54GM104944, GM103440).

<!--

Nomenclature/Notation

```{r ch070-all-dists, eval=FALSE}
all_dists <- list(
  list(dist = "beta", shape1, shape2),
  list(dist = "binom", size, prob),
  list(dist = "cauchy", location, scale),
  list(dist = "chisq", df),
  list(dist = "exp", rate),
  list(dist = "f", df1, df2),
  list(dist = "gamma", shape, rate),
  list(dist = "geom", prob),
  list(dist = "hyper", m, n, k),
  list(dist = "logis", location, scale),
  list(dist = "lnorm", meanlog, sdlog),
  list(dist = "nbinom", size, prob),
  list(dist = "norm", mean, sd),
  list(dist = "pois", lambda),
  list(dist = "t", df),
  list(dist = "unif", min, max),
  list(dist = "weibull", shape, scale),
  list(dist = "wilcox", m, n),
  list(dist = "signrank", n)
)
```

your comment -->