### 2021-03-01 010-background.Rmd
Revise 010 prior to CSDA submission

# Background {#background}

```{r ch010-background, include=FALSE, cache=FALSE}
box::use(
  dplyr[...],
  ggplot2[...],
  stringr[str_sub],
  GGally[ggpairs, wrap],
  knitr[kable]
)
load("data/example_genes.rda")
load("data/example_brca.rda")
d <- 1000
```

The `bigsimr` `R` package presented here provides multiple algorithms that operate with high-dimensional multivariate data; however, all these algorithms were originally designed to support a single task: to generate random vectors drawn from multivariate probability distributions with given marginal distributions and dependency metrics. Specifically, our goal is to efficiently simulate a large number, $B$, of random vectors ${\bf Y}=(Y_1, \ldots, Y_d)^\top$ with *correlated* components and heterogeneous marginal distributions, described via cumulative distribution functions (CDFs) $F_i$, where $d$ can be very large and still be computed in a practical timescale.

When designing this methodology, we developed the following properties to guide our effort. We divide the properties into two categories: (1) basic properties (BP) and "scalability" properties (SP). The BPs are adapted from an existing criteria due to @Nik13a. A suitable simulation strategy should possess the following properties:

* BP1: A wide range of dependences, allowing both positive and negative values, and, ideally, admitting the full range of possible values.
* BP2: Flexible dependence, meaning that the number of bivariate marginals can be equal to the number of dependence parameters.
* BP3: Flexible marginal modeling, generating heterogeneous data --- possibly from differing probability families.

Moreover, the simulation method must *scale* to high dimensions:

* SP1: Procedure must scale to high dimensions with practical compute times.
* SP2: Procedure must scale to high dimensions while maintaining accuracy.

<!-- 
The next section describes the motivating data set we use to provide example applications.
-->

## Motivating example: RNA-seq data

Simulating high-dimensional, non-normal, correlated data motivates this work --- in pursuit of modeling RNA-sequencing (RNA-seq) data [@Wang2009b; @Conesa2016b] derived from breast cancer patients. The RNA-seq data-generating process involves counting how often a particular form of human messenger RNA (mRNA) is expressed in a biological sample. RNA-seq platforms typically quantify the entire transcriptome in one experimental run, resulting in high-dimensional data. For human-derived samples, this results in count data corresponding to over 20,000 genes (protein-coding genomic regions) or even over 77,000 isoforms when alternatively-spliced mRNA are counted [@Schissler2019]. Importantly, due to inherent biological processes, gene expression data exhibit correlation --- co-expression --- across genes [@BE07; @Schissler2018].

We illustrate our methodology using a well-studied Breast Invasive Carcinoma (BRCA) data set housed in The Cancer Genome Atlas (TCGA; see Acknowledgments).
For simplicity, we only consider high expressing genes. In turn, we begin by filtering to retain the top `r d` of the highest-expressing genes (in terms of median expression) of the oover 20,000 gene measurements from $N=`r nrow(example_brca)`$ patients' tumor samples. This gives a great number of pairwise dependencies among the marginals (specifically, $`r choose(d,2)`$ correlation parameters). Table \@ref(tab:ch010-realDataTab) displays RNA-seq counts for three selected high-expressing genes for the first five patients' breast tumor samples.

```{r ch010-realDataTab}
set.seed(2020-02-25)
num_genes    <- 3
num_patients <- 5
gene_sample  <- sample(example_genes[1:1000], num_genes)

cap <- paste0("mRNA expression for three selected high-expressing genes, ",
              paste(gene_sample, collapse = ", "), 
              ", for the first five patients in the TCGA BRCA data set.")

small_brca <- example_brca %>%
  select(all_of(gene_sample)) %>%
  head(n = num_patients) %>%
  as_tibble(rownames = "Patient ID") %>%
  mutate(`Patient ID` = str_sub(`Patient ID`, end = 12))

small_brca %>%
  kable(booktabs = TRUE, caption = cap)
```

To help visualize the bivariate relationships for these three selected genes across all patients, Figure \@ref(fig:ch010-realDataFig) displays the marginal distributions and estimated Spearman's correlations (see Equation \@ref(eq:spearman) below).

```{r ch010-realDataFig, cache=FALSE, fig.width=8, fig.cap="Marginal scatterplots, densities, and estimated pairwise Spearman's correlations for three example genes from Table 1. The data possess heavy-right tails, are discrete, and have non-trivial intergene correlations. Modeling these data motivate our simulation methodology."}
ggpairs(
  data = example_brca[, gene_sample],
  upper = list(continuous = wrap('cor', method = "spearman"))
) + theme_bw()
```

## Measures of dependency

In multivariate analysis, an analyst must select a metric to quantify dependency.
The most widely-known is the Pearson (product-moment) correlation coefficient that describes the linear association between two random variables $X$ and $Y$, and, it is given by

\begin{equation}
(\#eq:pearson)
\rho_P(X,Y) = \frac{E(XY) - E(X)E(Y)}{\left[ \mathrm{Var}(X)\mathrm{Var}(Y)\right]^{1/2}}.
\end{equation}

As @MB13 and @MK01 discuss, for a bivariate normal $(X,Y)$ random vector, the Pearson correlation completely describes the dependency between the components. For non-normal marginals with monotone correlation patterns, however, $\rho_P$ suffers some drawbacks and may mislead or fail to capture important relationships [@MK01]. Alternatively in these settings, analysts often prefer rank-based correlation measures to describe the degree of monotonic association.

Two nonparametric, rank-based measures common in practice are Spearman's correlation (denoted $\rho_S$) and Kendall's $\tau$. Spearman's $\rho_S$ has an appealing correspondence as the Pearson correlation coefficient on *ranks* of the values, thereby captures nonlinear yet monotone relationships. Kendall's $\tau$, on the other hand, is the difference in probabilities of concordant and discordant pairs of observations, $(X_i, Y_i)$ and $(X_j, Y_j)$, with concordance meaning that orderings have the same direction (e.g., if $X_i < X_j$, then $Y_i < Y_j$). Note that concordance is determined by the ranks of the values, not the values themselves.

Both $\tau$ and $\rho_S$ are *invariant under monotone transformations* of the underlying random variates. As we will describe more fully in the [Algorithms](#algorithms) section, this property enables matching rank-based correlations with speed (SP1) and accuracy (SP2).

*Correspondence among Pearson, Spearman, and Kendall correlations*.
There is no closed form, general correspondence among the rank-based measures and the Pearson correlation coefficient, as the marginal distributions $F_i$ are intrinsic in their calculation. For *bivariate normal vectors*, however, the correspondence is well-known:

\begin{equation}
(\#eq:convertKendall)
\rho_{P} = \sin \left( \tau \times \frac{\pi}{2} \right), 
\end{equation}

\noindent and similarly for Spearman's $\rho$ [@K58],

\begin{equation}
(\#eq:convertSpearman)
\rho_P = 2 \times \sin \left( \rho_S \times \frac{\pi}{6} \right).
\end{equation}

*Discrete marginal considerations*.
Spearman's correlation for discrete marginal suffers issues whenever there is a nonzero probability of ties. For example, the Spearman's correlation of a discrete-valued random variable $X$ with itself could be less than 1 [@MB13]. One remedy for this issue is to rescale Equation \@ref(eq:spearman). For two random variables $X,Y$ with probability mass functions (PMFs) or probability densities functions (PDFs) $p(x)$ and $q(y)$, respectively, define the rescaled Spearman's correlation as

\begin{equation}
(\#eq:spearmanRescaled)
\rho_{RS}(X,Y) = \frac{\rho_s(X,Y)}{ \left[ \left[ 1 - \sum_x p(x)^3 \right] \left[ 1 - \sum_y q(y)^3 \right] \right]^{1/2}}.
\end{equation}

Note that with continuous marginals the rescaling returns $\rho_S$. For discrete marginals with large or infinite support, computing the adjustment factors $\sum_x p(x)^3$, $\sum_y p(y)^3$ over all large number of pairs becomes expensive (often violating desired property SP1). Further, the infinite sums must be approximated for count-valued data.

Also, for a discrete random variable $Y_i$, some care must be taken to define the quantile function $F_{i}^{-1}$. A standard practice is to set

\begin{equation}
F_{i}^{-1} = \inf\{y:F_{i}(y) \geq u \}.
(\#eq:inverseCDF)
\end{equation}

*Marginal-dependent bivariate correlation bounds*. 
Given two marginal distributions, $\rho_P$ is not free to vary over the entire range of possible correlations $[-1,1]$. The so-called *Frechet-Hoeffding bounds* are well-studied [@Nelsen2007; @BF17]. These constraints cannot be overcome through algorithm design. In general, the bounds are given by

\begin{equation}
(\#eq:frechet)
\rho_P^{max} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (U) \right), \quad \rho_P^{min} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (1 - U) \right)
\end{equation}

\noindent where $U$ is a uniform random variable on $(0,1)$ and $F^{-1}_1, F^{-1}_2$ are the inverse CDFs of $X_1$ and $X_2$, respectively, definitely by \@ref(eq:inverseCDF) when the variables are discrete.

## Gaussian copulas

There is a strong connection of our simulation strategy to Gaussian *copulas* (see @Nelsen2007 for a technical introduction). A copula is a distribution function on $[0,1]^d$ that describes a multivariate probability distribution with standard uniform marginals. This provides a powerful, natural way to characterize joint probability structures. Consequently, the study of copulas is an important and active area of statistical theory and practice.

For any random vector ${\bf X}=(X_1, \ldots, X_d)^\top$ with CDF $F$ and marginal CDFs $F_i$ there is a copula function $C(u_1, \ldots, u_d)$ satisfying

\begin{equation}
(\#eq:copula)
F(x_1, \ldots,x_d) = \mathbb P(X_1\leq x_1, \ldots,X_d\leq x_d) = C(F_1(x_1), \ldots, F_d(x_d)), 
\end{equation}

$x_i \in \mathbb R, i=1,\ldots,d.$

A Gaussian copula has marginal CDFs that are all standard normal, $F_i = \Phi, \forall \, i$. This representation corresponds to a multivariate normal (MVN) distribution with standard normal marginal distributions and covariance matrix ${\bf R_P}$. As the marginals are standardized to have unit variance, however, ${\bf R_P}$ is a Pearson correlation matrix. If $F_{{\bf R}}$ is the CDF of such a multivariate normal distribution, then the corresponding Gaussian copula $C_{{\bf R}}$ is defined through

\begin{equation}
(\#eq:gauss)
F_{{\bf R}}(x_1, \ldots, x_d) = C_{{\bf R}}(\Phi(x_1), \ldots, \Phi(x_d)),
\end{equation}

where $\Phi(\cdot)$ is the standard normal CDF. Note that the copula $C_{{\bf R}}$ is the familiar multivariate normal CDF of the random vector $(\Phi(X_1), \ldots, \Phi(X_d))$, where $(X_1, \ldots, X_d) \sim N_d({\bf 0}, {\bf R_P})$.

Sklar's Theorem [@Sklar1959; @Ubeda-Flores2017] guarantees that given inverse CDFs $F_i^{-1}$s and a valid correlation matrix (within the Frechet bounds) a random vector can be obtained via transformations involving copula functions. For example, using Gaussian copulas, we can construct a random vector ${\bf Y}  = (Y_1, \ldots,  Y_d)^\top$ with $Y_i \sim F_i$, viz. $Y_i = F_i^{-1}(\Phi(X_i)), i=1, \ldots, d$, where $(X_1, \ldots, X_d) \sim N_d({\bf 0}, {\bf R_P})$.