### 2021-03-01 030-bigsimr-package.Rmd

# The `bigsimr` R package {#package}

```{r ch030-basic-example-setup, echo=TRUE, cache=FALSE}
box::use(
  patchwork[...],
  dplyr[...],
  ggplot2[...],
  tidyr[drop_na],
  bigsimr[bigsimr_setup, distributions_setup],
  JuliaCall[julia_call]
)

Sys.setenv(JULIA_NUM_THREADS = parallel::detectCores())
bs <- bigsimr_setup(pkg_check = FALSE)
dist <- distributions_setup()
```

This section describes a low-dimensional (2D) random vector simulation workflow via the `bigsimr` `R` package (https://github.com/SchisslerGroup/r-bigsimr).
The package `bigsimr` provides an interface to the native code written in Julia, registered as the `Bigsimr` Julia package (https://github.com/adknudson/Bigsimr.jl).
In addition to the native Julia `Bigsimr` package and `R` interface `bigsimr`, we also provide a python interface `bigsimr` (https://github.com/SchisslerGroup/python-bigsimr/) that interfaces with the Julia `Bigsimr` package. 
The Julia implementation provides a high-performance implementation of our proposed random vector generation algorithm and associated functions (see Section \@ref(algorithms)).
When designing `bigsimr`, we aimed to conveniently provide parallelized computation, through multi-core and GPU acceleration, while allowing advanced users to customize.

The subsections below describe the basic use of `bigsimr`, by stepping through an example workflow using the data set `airquality` that contains daily air quality measurements in New York, May to September 1973 [@Chambers1983].
This workflow proceeds from computing environment setup, data wrangling, estimation, simulation configuration, random vector generation, and, finally, result visualization.
The code chunk below prepares the computing environment:

<!-- 
`bigsimr` is an R package for simulating high-dimensional multivariate data with arbitrary marginal distributions. The efficiency behind generating multivariate samples comes from the ability to utilize a GPU during the generation of multivariate normal samples and the subsequent transformation to uniform samples (see section \@ref(rand-vec-gen)). A GPU is not necessary in order to use `bigsimr`, and the speed benefits generally only come when simulating very high dimensional data ($d > 10000$).

The next computational step after obtaining the correlated uniform marginals is the inverse transform into the target marginal distributions. Since this consists of $d$ independent transformations, we utilize parallelization to achieve higher throughput. As with many parallel algorithms, there is overhead associated with forking the task to utilize multiple cores, however cost is negligible compared to the total run-time.
-->

## Bivariate example description

We illustrate the use of `bigsimr` motivated by the New York air quality data set (`airquality`) included in the R `datasets` package.
First, we load the `bigsimr` library and a few other convenient data science packages, including the syntactically-elegant `tidyverse` suite of `R` packages.

For simplicity and to provide a minimal working example, we consider bivariate simulation of the two `airquality` variables `Temperature`, in degrees Fahrenheit, and `Ozone` level, in parts per billion.

```{r ch030-air-quality-data, echo=TRUE}
df <- airquality %>%
  select(Temp, Ozone) %>%
  drop_na()
```

```{r ch030-aq-glimpse}
glimpse(df)
```

Figure \@ref(fig:ch030-aq-joint-dist) visualizes the bivariate relationship between Ozone and Temperature. We aim to simulate random two-component vectors mimicking this structure. The margins are not normally distributed; particularly the Ozone level exhibits a strong positive skew.

```{r ch030-aq-joint-dist, cache=FALSE, fig.width= 8, fig.cap="Bivariate scatterplot of Ozone vs. Temp with estimated marginal densities. We model the Ozone data as marginally log-normal and the Temperature data as normal."}
p0 <- ggplot(df, aes(Temp, Ozone)) +
  geom_point(size = 1) +
  theme(legend.position = "none") + 
  labs(x = "Temperature")

pTemp <- ggplot(df, aes(Temp)) + 
  geom_density() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

pOzone <- ggplot(df, aes(Ozone)) + 
    geom_density() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  coord_flip()

pTemp + plot_spacer() + p0 + pOzone + 
  plot_layout(widths = c(3,1), heights = c(1, 3)) 
```

Next, we specify the marginal distributions and correlation coefficient (both type and magnitude). Here the analyst is free to be creative. For this example, we avoid goodness-of-fit considerations to determine the marginal distributions. But it seems sensible without domain knowledge to estimate these quantities from the data and `bigsimr` contains fast functions designed for this task.

## Specifying marginal distributions

Based on the estimated densities in Figure \@ref(fig:ch030-aq-joint-dist), we assume `Temp` is normally distributed and `Ozone` is log-normally distributed, as the latter values are positive and skewed. We use the well-known, unbiased estimators for the normal distribution's parameters and maximum likelihood estimators for the log-normal parameters:

```{r ch030-aq-temp-pars, echo=TRUE}
df %>% 
  select(Temp) %>% 
  summarise_all(.funs = c(mean = mean, sd = sd))
```

```{r ch030-aq-ozone-pars, echo=TRUE}
mle_mean <- function(x) mean(log(x))
mle_sd <- function(x) mean( sqrt( (log(x) - mean(log(x)))^2 ) )

df %>% 
  select(Ozone) %>% 
  summarise_all(.funs = c(meanlog = mle_mean, sdlog = mle_sd))
```

Next, we configure the input marginals for later input into `rvec`. The marginal distributions are specified using `Julia`'s `Distributions` package and stored in a vector.

```{r ch030-margins-alist, echo=TRUE}
margins <- c(
  dist$Normal(mean(df$Temp), sd(df$Temp)),
  dist$LogNormal(mle_mean(df$Ozone), mle_sd(df$Ozone))
)
```

## Specifying correlation

As mentioned, the user must decide how to describe correlation, based on the particulars of the problem. For non-normal data and for improved simulation accuracy/scalability in our scheme, we advocate the use of Spearman's $\rho$ correlation matrix $R_S$ and Kendall's $\tau$ correlation matrix $R_K$. We also support Pearson correlation coefficient matching, while cautioning the user to check the performance for their parametric multivariate model (see [Monte Carlo evaluations](#simulations) below for evaluation strategies and guidance). These estimation methods are classical approaches, not designed for high-dimensional correlation estimation (see the [Conclusion and Discussion]({#discussion) sections for more on this).

```{r ch030-aq-cor, echo=TRUE}
(R_S <- bs$cor(as.matrix(df), bs$Spearman))
```

## Checking target correlation matrix admissibility

First we use `cor_bounds` to ensure that the pairwise target correlation values are valid prior to mapping step which constructs $R_X$ for the MVN input into NORTA (see Section \@ref(algorithms)). `cor_bounds` estimates the pairwise lower and upper correlation bounds using the Generate, Sort, and Correlate algorithm of @DH2011.

```{r ch030-cor-bounds, echo=TRUE}
bounds <- bs$cor_bounds(margins)
bounds$lower
```

Since our single estimated Spearman correlation is within the theoretical bounds, the correlation is valid as input to `rvec`. But even if the 2-dimensional bounds are satisfied for each pair of margins, this does not guarantee the feasibility of a $d-$variate distribution [@BF17].

To provide higher dimensional feasibility/admissibility checking, we begin by mapping using `cor_convert` and check admissibility.

```{r ch030-cor-convert, echo=TRUE}
# Step 1. Mapping
(R_X <- bs$cor_convert(R_S, bs$Spearman, bs$Pearson))

# Step 2. Check admissibility
julia_call("Bigsimr.iscorrelation", R_X)
```

The bounds on the Pearson correlation coefficient $R_P$ between these margins are restricted as seen below in our MC estimated correlation bounds:

```{r ch030-cor-bounds-pearson, echo=TRUE}
bs$cor_bounds(margins[1], margins[2], bs$Pearson, n_samples = 1e6)
```

Our MC estimate of the bounds slightly underestimates the theoretic bounds of $(-0.881, 0.881)$. An analytic derivation is presented as an Appendix.

## Simulating random vectors
Finally, we arrive at the main function of `bigsimr`, `rvec`. We now simulate $B=10,000$ random vectors from the assumed joint distribution of Ozone levels and Temp.

```{r ch030-sim-margins, echo=TRUE}
x <- bs$rvec(10000, R_X, margins)
df_sim <- as.data.frame(x)
colnames(df_sim) <- colnames(df)
```

Figure \@ref(fig:ch030-plot-sim) plots the 10,000 simulated points.

```{r ch030-plot-sim, fig.cap="Contour plot and marginal densities for the simulated bivariate distribution of Air Quality Temperatures and Ozone levels. The simulated points mimic the observed data with respect to both the marginal characteristics and bivariate association."}
p1 <- df_sim %>%
  ggplot(aes(Temp, Ozone)) +
  geom_density_2d_filled() +
  theme(legend.position = "none") +
  labs(x = "Simulated Temperature", y = "Simulated Ozone") +
  scale_y_continuous(limits = c(0, max(df$Ozone))) +
  scale_x_continuous(limits = range(df$Temp))

p1Temp <- ggplot(df_sim, aes(Temp)) + 
  geom_density(alpha = 0.5, fill = "lightseagreen") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

p1Ozone <- ggplot(df_sim, aes(Ozone)) + 
  geom_density(alpha = 0.5, fill = "lightseagreen") + 
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  coord_flip()

p1Temp + plot_spacer() + p1 + p1Ozone + 
  plot_layout(widths = c(3,1), heights = c(1, 3))
```