### 2021-04-01 article_bigsimr.Rmd

---
documentclass: jss
author:
  - name: Alfred G. Schissler
    affiliation: University of Nevada
    address: |
      | 1664 N Virginia St.
      | Reno, NV 89557
    email: \email{aschissler@unr.edu}
  - name: Edward J. Bedrick
    affiliation: University of Nevada
  - name: Alexander D. Knudson
    affiliation: University of Nevada \AND
  - name: Tomasz J. Kozubowski
    affiliation: University of Nevada
  - name: Tin Nguyen
    affiliation: University of Nevada
  - name: Anna K. Panorska
    affiliation: University of Nevada \AND
  - name: Juli Petereit
    affiliation: University of Nevada
  - name: Walter W. Piegorsch
    affiliation: University of Nevada
  - name: Duc Tran
    affiliation: University of Nevada
title:
  formatted: "Simulating High-Dimensional Multivariate Data using the \\pkg{Bigsimr} Package"
  plain:     "Simulating High-Dimensional Multivariate Data using the Bigsimr Package"
  short:     "\\pkg{Bigsimr}: Simulate High-Dimensional Data"
abstract: >
  It is critical to realistically simulate data when employing Monte Carlo techniques and evaluating statistical methodology. Measurements are often correlated and high dimensional in this era of big data, such as data obtained through high-throughput biomedical experiments. Due to computational complexity and a lack of user-friendly software available to simulate these massive multivariate constructions, researchers resort to simulation designs that posit independence or perform arbitrary data transformations. This article introduces the \\pkg{Bigsimr} R package. This high-level package provides a flexible and scalable procedure to simulate high-dimensional random vectors with arbitrary marginal distributions and known Pearson, Spearman, or Kendall correlation matrix. \\pkg{Bigsimr} contains additional high-performance features, including multi-core algorithms to simulate random vectors, estimate correlation, and compute the nearest correlation matrix. Monte Carlo studies quantify the accuracy and scalability of our approach, up to $d=10,000$. Finally, we demonstrate example applications enabled via \\pkg{Bigsimr} by applying these methods to a motivating dataset --- RNA-sequencing data obtained from breast cancer tumor samples.
keywords:
  formatted: [multivariate simulation, high-dimensional data, nonparametric correlation, Gaussian copula, RNA-sequencing data, breast cancer]
  plain:     [multivariate simulation, high-dimensional data, nonparametric correlation, Gaussian copula, RNA-sequencing data, breast cancer]
bibliography: [bigsimr.bib, packages.bib, alex.bib]
biblio-style: jss
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{amsfonts}
  \usepackage{amsthm}
  \newtheorem{theorem}{Theorem}
  \newtheorem{lemma}[theorem]{Lemma}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{ragged2e}
  \usepackage{setspace}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')

options(prompt = 'R> ', 
        continue = '+ ',
        digits = 4)

get_rmd <- function(str_no = "") list.files(path = ".", pattern = str_no)

knitr::opts_chunk$set(
  # Text Results
  eval = TRUE,
  echo = FALSE,
  results = 'markup',
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  # Code Decoration
  comment = NA,
  background = '#F7F7F7',
  # Cache
  cache = TRUE,
  # Plots
  fig.show = "hold",
  fig.width = 6.5,
  fig.asp = 0.618,  # 1 / phi
  out.width = "85%",
  fig.align = 'center'
)

# options(dplyr.print_min = 6, dplyr.print_max = 6)

seed_number <- 2021-03-23
set.seed(seed = seed_number)
```

<!-- New page after title and abstract -->
\newpage

<!-- Set double spaces -->
\setstretch{2.0}

<!-- Main content -->

```{r, child=get_rmd("000")}
```

```{r, child=get_rmd("010")}
```

```{r, child=get_rmd("020")}
```

```{r, child=get_rmd("030")}
```

```{r, child=get_rmd("040")}
```

```{r, child=get_rmd("050")}
```

```{r, child=get_rmd("060")}
```

<!-- Supplementary materials -->
\newpage

```{r, child=get_rmd("070")}
```

<!-- Appendix -->
\newpage

```{r, child=get_rmd("080")}
```

<!-- Turn off double spaces -->
\setstretch{1.0}