### 2021-04-01 article_bigsimr.tex

\documentclass[
]{jss}

\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Alfred G. Schissler\\University of Nevada \And Edward J.
Bedrick\\University of Nevada \And Alexander D. Knudson\\University of
Nevada \AND Tomasz J. Kozubowski\\University of Nevada \And Tin
Nguyen\\University of Nevada \And Anna K. Panorska\\University of Nevada
\AND Juli Petereit\\University of Nevada \And Walter W.
Piegorsch\\University of Nevada \And Duc Tran\\University of Nevada
}
\title{Simulating High-Dimensional Multivariate Data using the
\pkg{Bigsimr} Package}

\Plainauthor{Alfred G. Schissler, Edward J. Bedrick, Alexander D.
Knudson, Tomasz J. Kozubowski, Tin Nguyen, Anna K. Panorska, Juli
Petereit, Walter W. Piegorsch, Duc Tran}
\Plaintitle{Simulating High-Dimensional Multivariate Data using the
Bigsimr Package}
\Shorttitle{\pkg{Bigsimr}: Simulate High-Dimensional Data}

\Abstract{
It is critical to realistically simulate data when employing Monte Carlo
techniques and evaluating statistical methodology. Measurements are
often correlated and high dimensional in this era of big data, such as
data obtained through high-throughput biomedical experiments. Due to
computational complexity and a lack of user-friendly software available
to simulate these massive multivariate constructions, researchers resort
to simulation designs that posit independence or perform arbitrary data
transformations. This article introduces the
\textbackslash pkg\{Bigsimr\} R package. This high-level package
provides a flexible and scalable procedure to simulate high-dimensional
random vectors with arbitrary marginal distributions and known Pearson,
Spearman, or Kendall correlation matrix. \textbackslash pkg\{Bigsimr\}
contains additional high-performance features, including multi-core
algorithms to simulate random vectors, estimate correlation, and compute
the nearest correlation matrix. Monte Carlo studies quantify the
accuracy and scalability of our approach, up to \(d=10,000\). Finally,
we demonstrate example applications enabled via
\textbackslash pkg\{Bigsimr\} by applying these methods to a motivating
dataset --- RNA-sequencing data obtained from breast cancer tumor
samples.
}

\Keywords{multivariate simulation, high-dimensional data, nonparametric
correlation, Gaussian copula, RNA-sequencing data, breast cancer}
\Plainkeywords{multivariate simulation, high-dimensional
data, nonparametric correlation, Gaussian copula, RNA-sequencing
data, breast cancer}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    Alfred G. Schissler\\
    University of Nevada\\
    1664 N Virginia St.\\
Reno, NV 89557\\
  E-mail: \email{aschissler@unr.edu}\\

                  }

% Pandoc citation processing

% Pandoc header

\usepackage{amsmath} \usepackage{amssymb} \usepackage{amsfonts} \usepackage{amsthm} \newtheorem{theorem}{Theorem} \newtheorem{lemma}[theorem]{Lemma} \usepackage{booktabs} \usepackage{longtable} \usepackage{array} \usepackage{ragged2e} \usepackage{setspace}

\begin{document}

\newpage

\setstretch{2.0}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Massive high-dimensional (HD) data sets are now commonplace in many
areas of scientific inquiry. As new methods are developed for data, a
fundamental challenge lies in designing and conducting simulation
studies to assess the operating characteristics of analyzing such
proposed methodology --- such as false positive rates, statistical
power, interval coverage, and robustness --- often with comparison to
existing methods. Further, efficient simulation empowers statistical
computing strategies, such as the parametric bootstrap
\citep{Chernick2008} to simulate from a hypothesized null model,
providing inference in analytically challenging settings. Such Monte
Carlo (MC) techniques become difficult for high-dimensional data using
existing algorithms and tools. This is particularly true when simulating
massive multivariate, non-normal distributions, arising naturally in
many fields of study.

As many have noted, it can be vexing to simulate dependent,
non-normal/discrete data, even for low dimensional settings
\citep{MB13, XZ19}. For continuous non-normal multivariate data, the
well-known NORmal To Anything (NORTA) algorithm \citep{Cario1997} and
other copula approaches \citep{Nelsen2007} are well-studied, with
flexible, robust software available \citep{Yan2007, Chen2001}. Yet these
approaches do not scale in a timely fashion to high-dimensional problems
\citep{Li2019gpu}. For discrete data, early simulation strategies had
major flaws, such as failing to obtain the full range of possible
correlations (e.g., admitting only positive correlations: see
\citet{Park1996}). While more recent approaches
\citep{MB13, Xia17, BF17} have largely remedied this issue for
low-dimensional problems, the existing tools are not designed to scale
to high dimensions.

Another central issue lies in characterizing dependence between
components in the high-dimensional random vector. The choice of
correlation in practice usually relates to the eventual analytic goal
and distributional assumptions of the data (e.g., non-normal, discrete,
infinite support, etc). For normal data, the Pearson product-moment
correlation describes the dependency perfectly. As we will see, however,
simulating arbitrary random vectors that match a target Pearson
correlation matrix is computationally intense \citep{Chen2001, Xia17}.
On the other hand, an analyst might consider use of nonparametric
correlation measures to better characterize monotone, non-linear
dependence, such as Spearman's \(\rho\) and Kendall's \(\tau\).
Throughout, we focus on matching these nonparametric dependence
measures, as our aim lies in modeling non-normal data and these
rank-based measures possess invariance properties favorable in our
proposed methodology. We do, however, provide Pearson matching with some
caveats.

With all this in mind, we present a scalable, flexible multivariate
simulation algorithm. The crux of the method lies in the construction of
a Gaussian copula in the spirit of the NORTA procedure. Further, we
introduce the \texttt{bigsimr} R package that provides high-performance
software implementing our algorithm. The algorithm design leverages
useful properties of nonparametric correlation measures, namely
invariance under monotone transformation and well-known closed-form
relationships between dependence measures for the multivariate normal
(MVN) distribution.

Our study proceeds by providing background information, including a
description of a motivating example application: RNA-sequencing
(RNA-seq) breast cancer data. Then we describe and justify our
simulation methodology and related algorithms. Next, we detail an
illustrative low-dimensional example of basic use of the
\texttt{bigsimr} R package. Then we proceed with Monte Carlo studies
under various bivariate distributional assumptions to assess accuracy.
We conclude the Monte Carlo evaluations by summarizing the computation
time for increasingly higher dimensional vectors. After the MC
evaluations, we simulate random vectors motivated by our RNA-seq
example, evaluate the accuracy, and provide example statistical
computing tasks, namely MC estimation of joint probabilities and
evaluating HD correlation estimation efficiency. Finally, we discuss the
method's utility, limitations, and future directions.

\hypertarget{background}{%
\section{Background}\label{background}}

The \texttt{bigsimr} \texttt{R} package presented here provides multiple
algorithms that operate with high-dimensional multivariate data;
however, all these algorithms were originally designed to support a
single task: to generate random vectors drawn from multivariate
probability distributions with given marginal distributions and
dependency metrics. Specifically, our goal is to efficiently simulate a
large number, \(B\), of HD random vectors
\({\bf Y}=(Y_1, \ldots, Y_d)^\top\) with \emph{correlated} components
and heterogeneous marginal distributions, described via cumulative
distribution functions (CDFs) \(F_i\).

When designing this methodology, we developed the following properties
to guide our effort. We divide the properties into two categories: (1)
basic properties (BP) and ``scalability'' properties (SP). The BPs are
adapted from an existing criteria due to \citet{Nik13a}. A suitable
simulation strategy should possess the following properties:

\setstretch{1.5}

\begin{itemize}
\tightlist
\item
  BP1: A wide range of dependences, allowing both positive and negative
  values, and, ideally, admitting the full range of possible values.
\item
  BP2: Flexible dependence, meaning that the number of bivariate
  marginals can be equal to the number of dependence parameters.
\item
  BP3: Flexible marginal modeling, generating heterogeneous data ---
  possibly from differing probability families.
\end{itemize}

\setstretch{2.0}

Moreover, the simulation method must \emph{scale} to high dimensions:

\setstretch{1.5}

\begin{itemize}
\tightlist
\item
  SP1: Procedure must scale to high dimensions with practical compute
  times.
\item
  SP2: Procedure must scale to high dimensions while maintaining
  accuracy.
\end{itemize}

\setstretch{2.0}

\hypertarget{motivating-example-rna-seq-data}{%
\subsection{Motivating example: RNA-seq
data}\label{motivating-example-rna-seq-data}}

Simulating high-dimensional, non-normal, correlated data motivates this
work --- in pursuit of modeling RNA-sequencing (RNA-seq) data
\citep{Wang2009b, Conesa2016b} derived from breast cancer patients. The
RNA-seq data-generating process involves counting how often a particular
form of human messenger RNA (mRNA) is expressed in a biological sample.
RNA-seq platforms typically quantify the entire transcriptome in one
experimental run, resulting in high-dimensional data. For human-derived
samples, this results in count data corresponding to over 20,000 genes
(protein-coding genomic regions) or even over 77,000 isoforms when
alternatively-spliced mRNA are counted \citep{Schissler2019}.
Importantly, due to inherent biological processes, gene expression data
exhibit correlation (co-expression) across genes
\citep{BE07, Schissler2018}.

We illustrate our methodology using a well-studied Breast Invasive
Carcinoma (BRCA) data set housed in The Cancer Genome Atlas (TCGA; see
Acknowledgments). For ease of modeling and simplicity of exposition, we
only consider high expressing genes. In turn, we begin by filtering to
retain the top 1000 of the highest-expressing genes (in terms of median
expression) of the over 20,000 gene measurements from \(N=878\)
patients' tumor samples. This gives a great number of pairwise
dependencies among the marginals (specifically,
\(\ensuremath{4.995\times 10^{5}}\) correlation parameters). Table
@ref(tab:ch010-realDataTab) displays RNA-seq counts for three selected
high-expressing genes for the first five patients' breast tumor samples.
To help visualize the bivariate relationships for these three selected
genes across all patients, Figure @ref(fig:ch010-realDataFig) displays
the marginal distributions and estimated Spearman's correlations.

\begin{CodeChunk}
\begin{table}

\caption{\label{tab:ch010-realDataTab}mRNA expression for three selected high-expressing genes, STAU1, FKBP1A, NME2, for the first five patients in the TCGA BRCA data set.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Patient ID & STAU1 & FKBP1A & NME2\\
\midrule
TCGA-A1-A0SB & 10440 & 11354 & 17655\\
TCGA-A1-A0SD & 21523 & 20221 & 14653\\
TCGA-A1-A0SE & 21733 & 22937 & 35251\\
TCGA-A1-A0SF & 11866 & 19650 & 16551\\
TCGA-A1-A0SG & 12486 & 12089 & 10434\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{article_bigsimr_files/figure-latex/ch010-realDataFig-1}

}

\caption[Marginal scatterplots, densities, and estimated pairwise Spearman's correlations for three example genes from Table 1]{Marginal scatterplots, densities, and estimated pairwise Spearman's correlations for three example genes from Table 1. The data possess outliers, heavy-right tails, are discrete, and have non-trivial intergene correlations. Modeling these data motivate our simulation methodology.}\label{fig:ch010-realDataFig}
\end{figure}
\end{CodeChunk}

\hypertarget{measures-of-dependency}{%
\subsection{Measures of dependency}\label{measures-of-dependency}}

In multivariate analysis, an analyst must select a metric to quantify
dependency. The most widely-known is the Pearson (product-moment)
correlation coefficient that describes the linear association between
two random variables \(X\) and \(Y\), and, it is given by

\begin{equation}
\rho_P(X,Y) = \frac{E(XY) - E(X)E(Y)}{\left[ \mathrm{Var}(X)\mathrm{Var}(Y)\right]^{1/2}}.
(\#eq:pearson)
\end{equation}

As \citet{MB13} and \citet{MK01} discuss, for a bivariate normal
\((X,Y)\) random vector, the Pearson correlation completely describes
the dependency between the components. For non-normal marginals with
monotone correlation patterns, however, \(\rho_P\) suffers some
drawbacks and may mislead or fail to capture important relationships
\citep{MK01}. Alternatively in these settings, analysts often prefer
rank-based correlation measures to describe the degree of monotonic
association.

Two nonparametric, rank-based measures common in practice are Spearman's
correlation (denoted \(\rho_S\)) and Kendall's \(\tau\). Spearman's
\(\rho_S\) has an appealing correspondence as the Pearson correlation
coefficient on \emph{ranks} of the values, thereby captures nonlinear
yet monotone relationships. Kendall's \(\tau\), on the other hand, is
the difference in probabilities of concordant and discordant pairs of
observations, \((X_i, Y_i)\) and \((X_j, Y_j)\), with concordance
meaning that orderings have the same direction (e.g., if \(X_i < X_j\),
then \(Y_i < Y_j\)). Note that concordance is determined by the ranks of
the values, not the values themselves.

Both \(\tau\) and \(\rho_S\) are \emph{invariant under monotone
transformations} of the underlying random variates. As we will describe
more fully in the \protect\hyperlink{algorithms}{Algorithms} section,
this property enables matching rank-based correlations with speed (SP1)
and accuracy (SP2).

\emph{Correspondence among Pearson, Spearman, and Kendall correlations}

There is no closed form, general correspondence among the rank-based
measures and the Pearson correlation coefficient, as the marginal
distributions \(F_i\) are intrinsic in their calculation. For
\emph{bivariate normal vectors}, however, the correspondence is
well-known:

\begin{equation}
(\#eq:convertKendall)
\rho_{P} = \sin \left( \tau \times \frac{\pi}{2} \right), 
\end{equation}

\noindent and similarly for Spearman's \(\rho\) \citep{K58},

\begin{equation}
(\#eq:convertSpearman)
\rho_P = 2 \times \sin \left( \rho_S \times \frac{\pi}{6} \right).
\end{equation}

\emph{Marginal-dependent bivariate correlation bounds}

Given two marginal distributions, \(\rho_P\) is not free to vary over
the entire range of possible correlations \([-1,1]\). The so-called
\emph{Frechet-Hoeffding bounds} are well-studied
\citep{Nelsen2007, BF17}. These constraints cannot be overcome through
algorithm design. In general, the bounds are given by

\begin{equation}
(\#eq:frechet)
\rho_P^{max} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (U) \right), \quad \rho_P^{min} = \rho_P \left( F^{-1}_1 (U), F^{-1}_2 (1 - U) \right)
\end{equation}

\noindent where \(U\) is a uniform random variable on \((0,1)\) and
\(F^{-1}_1, F^{-1}_2\) are the inverse CDFs of \(X_1\) and \(X_2\),
respectively, definitely by @ref(eq:inverseCDF) when the variables are
discrete.

\begin{equation}
F_{i}^{-1} = \inf\{y:F_{i}(y) \geq u \}.
(\#eq:inverseCDF)
\end{equation}

\hypertarget{gaussian-copulas}{%
\subsection{Gaussian copulas}\label{gaussian-copulas}}

There is a strong connection of our simulation strategy to Gaussian
\emph{copulas} (see \citet{Nelsen2007} for a technical introduction). A
copula is a distribution function on \([0,1]^d\) that describes a
multivariate probability distribution with standard uniform marginals.
This provides a powerful, natural way to characterize joint probability
structures. Consequently, the study of copulas is an important and
active area of statistical theory and practice.

For any random vector \({\bf X}=(X_1, \ldots, X_d)^\top\) with CDF \(F\)
and marginal CDFs \(F_i\) there is a copula function
\(C(u_1, \ldots, u_d)\) satisfying

\begin{equation}
F(x_1, \ldots, x_d) = {\mathbb P}(X_1\leq x_1, \ldots,X_d\leq x_d) = C(F_1(x_1), \ldots, F_d(x_d)), 
(\#eq:copula)
\end{equation}

\(x_i \in {\mathbb R}, i=1,\ldots,d.\)

A Gaussian copula has marginal CDFs that are all standard normal,
\(F_i = \Phi, \forall \, i\). This representation corresponds to a
multivariate normal (MVN) distribution with standard normal marginal
distributions and covariance matrix \({\bf R_P}\). As the marginals are
standardized to have unit variance, however, \({\bf R_P}\) is a Pearson
correlation matrix. If \(F_{{\bf R}}\) is the CDF of such a multivariate
normal distribution, then the corresponding Gaussian copula
\(C_{{\bf R}}\) is defined through

\begin{equation}
(\#eq:gauss)
F_{{\bf R}}(x_1, \ldots, x_d) = C_{{\bf R}}(\Phi(x_1), \ldots, \Phi(x_d)),
\end{equation}

where \(\Phi(\cdot)\) is the standard normal CDF. Note that the copula
\(C_{{\bf R}}\) is the familiar multivariate normal CDF of the random
vector \((\Phi(X_1), \ldots, \Phi(X_d))\), where
\((X_1, \ldots, X_d) \sim N_d({\bf 0}, {\bf R_P})\).

Sklar's Theorem \citep{Sklar1959, Ubeda-Flores2017} guarantees that
given inverse CDFs \(F_i^{-1}\)s and a valid correlation matrix (within
the Frechet bounds) a random vector can be obtained via transformations
involving copula functions. For example, using Gaussian copulas, we can
construct a random vector \({\bf Y} = (Y_1, \ldots, Y_d)^\top\) with
\(Y_i \sim F_i\), viz.~\(Y_i = F_i^{-1}(\Phi(X_i)), i=1, \ldots, d\),
where \((X_1, \ldots, X_d) \sim N_d({\bf 0}, {\bf R_P})\).

\hypertarget{algorithms}{%
\section{Algorithms}\label{algorithms}}

This section describes our methods involved in simulating a random
vector \(\bf Y\) with \(Y_i\) components for \(i=1,\ldots,d\). Each
\(Y_i\) has a specified marginal CDF \(F_i\) and its inverse
\(F^{-1}_i\). To characterize dependency, every pair \((Y_i, Y_j)\) has
a given Pearson correlation \(\rho_P\), Spearman correlation \(\rho_S\),
and/or Kendall's \(\tau\). The method can be described as a
\emph{high-performance Gaussian copula} (Equation @ref(eq:gauss))
providing a high-dimensional NORTA-inspired algorithm.

\hypertarget{normal-to-anything-norta}{%
\subsection{NORmal To Anything (NORTA)}\label{normal-to-anything-norta}}

The well-known NORTA algorithm \citep{Cario1997} simulates a random
vector \(\bf Y\) with variance-covariance matrix \(\Sigma_{\bf Y}\).
Specifically, NORTA algorithm proceeds as:

\setstretch{1.5}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Simulate a random vector \(\bf Z\) with \(d\) \emph{independent and
  identically distributed} (iid) standard normal components.
\item
  Determine the input matrix \(\Sigma_{\bf Z}\) that corresponds with
  the specified output \(\Sigma_{\bf Y}\).
\item
  Produce a Cholesky factor \(M\) of \(\Sigma_{\bf Z}\) such that
  \(M M^{\prime}=\Sigma_{\bf Z}\).
\item
  Set \(X\) by \(X \gets MZ\).
\item
  \(\text{Return} \; Y \; \text{where} \; Y_i \gets F_i^{-1}[\Phi(X_i)], \; i=1,...,d\).
\end{enumerate}

\setstretch{2.0}

With modern parallel computing, steps 1, 3, 4, 5 are readily implemented
as high-performance, multi-core and/or graphical-processing-unit (GPU)
accelerated algorithms --- providing fast scalability using
readily-available hardware.

Matching specified Pearson correlation coefficients exactly (step 2
above), however, is computationally costly. In general, there is no
closed-form correspondence between the components of the input
\(\Sigma_{\bf Z}\) and target \(\Sigma_{\bf Y}\). Matching the
correlations involves evaluating or approximating \(\binom{d}{2}\)
integrals of the form

\begin{equation}
    \mathrm{E}\left[Y_i Y_j\right] = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} F_i^{-1}\left[\Phi(z_i)\right] F_j^{-1}\left[\Phi(z_j)\right] \phi(z_i, z_j, \rho_z) dz_i dz_j,
    (\#eq:pearsonIntegralRelation)
\end{equation}

where \(\phi(\cdot)\) is the joint probability distribution function of
two correlated standard normal variables. For HD data, (nearly) exact
evaluation may become too costly to enable practical simulation studies.
For low-dimensional problems, however, methods and tools exist to match
Pearson correlations precisely; see \citep{Xia17} and the publicly
available \texttt{nortaRA} R package \citep{Chen2001}. As described
later, to enable HD Pearson matching we approximate these integrals.

\emph{NORTA in higher dimensions}

Sklar's theorem provides a useful characterization of multivariate
distributions through copulas. Yet the choice of copula-based simulation
algorithm affects which joint distributions may be simulated. Even in
low-dimensional spaces (e.g., \(d=3\)), there exist valid multivariate
distributions with \emph{feasible} Pearson correlation matrices that
NORTA cannot match exactly \citep{LH75}. This occurs when the bivariate
transformations are applied to find the input correlation matrix, yet
when combined the resultant matrix is indefinite. These situations do
occur, even using exact analytic calculations. Such problematic target
correlation matrices are termed \emph{NORTA defective}.

\citet{GH02} conducted a Monte Carlo study to estimate the probability
of encountering NORTA defective matrices while increasing the dimension
\(d\). They found that for what is now considered low-to-moderate
dimensions (\(d \approx 20\)), almost \emph{all} feasible matrices are
NORTA defective. This stems from the concentration of measure near the
boundary of the space of all possible correlation matrices as dimension
increases. Unfortunately, it is precisely near this boundary that NORTA
defective matrices reside.

There is hope, however, as \citet{GH02} also showed that replacing an
indefinite input correlation matrix with a close proxy will give
approximate matching to the target --- with adequate performance for
moderate \(d\). This provides evidence that our nearest positive
definite (PD) augmented approach will maintain reasonable accuracy if
our input matching scheme returns an indefinite matrix, at least for the
rank-based matching scheme described above.

\hypertarget{rand-vec-gen}{%
\subsection{Random vector generator}\label{rand-vec-gen}}

We now describe our algorithm to generate random vectors, which mirrors
the classical NORTA algorithm above with some modifications for
rank-based dependency matching:

\setstretch{1.5}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Mapping step

  \begin{itemize}
  \item
    \begin{enumerate}
    \def\labelenumii{(\roman{enumii})}
    \tightlist
    \item
      Convert the target Spearman correlation matrix \(R_S\) to the
      corresponding MVN Pearson correlation \(R_X\). Alternatively,
    \end{enumerate}
  \item
    (i') Convert the target Kendall \(\tau\) matrix \(R_K\) to the
    corresponding MVN Pearson correlation \(R_X\). Alternatively,
  \item
    (i'\,') Convert the target Pearson correlation matrix to \(R_P\) to
    the corresponding, approximate MVN Pearson correlation \(R_X\).
  \end{itemize}
\item
  Check admissibility and, if needed, compute the nearest correlation
  matrix.

  \begin{itemize}
  \item
    \begin{enumerate}
    \def\labelenumii{(\roman{enumii})}
    \tightlist
    \item
      Check that \(R_X\) is a correlation matrix, a positive definite
      matrix with 1's along the diagonal.
    \end{enumerate}
  \item
    \begin{enumerate}
    \def\labelenumii{(\roman{enumii})}
    \setcounter{enumii}{1}
    \tightlist
    \item
      If \(R_X\) is a correlation matrix, the input matrix is
      \emph{admissible} in this scheme. Otherwise
    \end{enumerate}
  \item
    (ii') Replace \(R_X\) with the nearest correlation matrix
    \(\tilde{R}_X\), in the Frobenius norm.
  \end{itemize}
\item
  Gaussian copula

  \begin{itemize}
  \item
    \begin{enumerate}
    \def\labelenumii{(\roman{enumii})}
    \tightlist
    \item
      Generate \({\bf X}=(X_1, \ldots, X_d) \sim N_d({\bf 0}, R_X)\).
    \end{enumerate}
  \item
    \begin{enumerate}
    \def\labelenumii{(\roman{enumii})}
    \setcounter{enumii}{1}
    \tightlist
    \item
      Transform \({\bf X}\) to \({\bf U} = (U_1, \ldots, U_d)\)
      viz.~\(U_i=\Phi(X_i)\), \(i=1, \ldots, d\).
    \end{enumerate}
  \item
    \begin{enumerate}
    \def\labelenumii{(\roman{enumii})}
    \setcounter{enumii}{2}
    \tightlist
    \item
      Return \({\bf Y} = (Y_1, \ldots, Y_d)\), where
      \(Y_i=F_i^{-1}(U_i)\), \(i=1, \ldots, d\).
    \end{enumerate}
  \end{itemize}
\end{enumerate}

\setstretch{2.0}

\emph{Step 1}

The first two descriptions of the \emph{Mapping step} employ the
closed-form relationships between \(\rho_S\) and \(\tau\) with
\(\rho_P\) for bivariate normal random variables via Equations
@ref(eq:convertKendall) and @ref(eq:convertSpearman), respectively
(implemented as \texttt{cor\_covert}). Initializing our algorithm to
match the nonparametric correlations by computing these equations for
all pairs is computationally trivial.

For the computationally expensive process to match Pearson correlations,
we approximate Equation @ref(eq:pearsonIntegralRelation) for all pairs
of margins. To this end, we implement the approximation scheme
introduced by \citet{xiao2019matching}. Briefly, the many double
integrals of the form in @ref(eq:pearsonIntegralRelation) are
approximated by weighted sums of Hermite polynomials. Matching
coefficients for pairs of continuous distributions is made tractable by
this method, but for discrete distributions (especially discrete
distributions with large support sets or infinite support), the problem
is not so simple or efficient.

Our solution is to approximate discrete distributions by a continuous
distribution. The question then becomes that of which distribution to
use. We found that Generalized S-Distributions \citep{muino2006gs} solve
this problem by approximating a wide range of unimodal distributions,
both continuous and discrete. Using the GS-Distribution approximation of
discrete distributions in Pearson matching scheme above yields favorable
results.

\emph{Step 2}

Once \(R_X\) has been determined, we check admissibility of the adjusted
correlation via the steps described above. If \(R_X\) is not a valid
correlation matrix, then we compute the nearest correlation matrix.
Finding the nearest correlation matrix is a common statistical computing
problem. The defacto default function in \texttt{R} is
\texttt{Matrix::nearPD}, an alternating projection algorithm due to
\citet{higham2002computing}. As implemented the function fails to scale
to HD. Instead, we provide the quadratically convergent algorithm based
on the theory of strongly semi-smooth matrix functions
(\citet{qi2006quadratically}). The nearest correlation matrix problem
can be written down as the following convex optimization problem:

\begin{align*}
    \mathrm{min}\quad & \frac{1}{2} \Vert G - X \Vert^2 \\
    \mathrm{s.t.}\quad & X_{ii} = 1, \quad i = 1, \ldots , n, \\
    & X \in S_{+}^{n}
\end{align*}

For nonparametric correlation measures, our algorithm allows the
generation of high-dimensional multivariate data with arbitrary marginal
distributions with a broad class of admissible Spearman correlation
matrices and Kendall \(\tau\) matrices. The admissible classes consist
of the matrices that map to a Pearson correlation matrix for a MVN. In
particular, if we let \(X\) be MVN with \(d\) components and set

\begin{align*}
\Omega_P &= \{ R_P : R_P \textrm{ is a correlation matrix for } X \} \\
\Omega_K &= \{ R_K : R_K \textrm{ is a correlation matrix for } X \} \\
\Omega_S &= \{ R_S : R_S \textrm{ is a correlation matrix for } X \} \\
\end{align*}

then there are 1-1 mappings between these sets. We conjecture informally
that the sets of admissible \(R_S\) and \(R_K\) are not highly
restrictive. In particular, \(R_P\) is approximately \(R_S\) for a MVN,
suggesting that the admissible set \(\Omega_S\) should be flexible as
\(R_P\) can be any PD matrix with 1's along the diagonal. We provide
methods to check whether a target \(R_S\) is an element of the
\emph{admissible set} \(\Omega_S\) (and similarly for \(R_K\)).

There is an increasing probability of encountering a non-admissible
correlation matrix as dimension increases. In our experience, the
mapping step for large \(d\) almost always produces a \(R_X\) that is
not a correlation matrix. In Section @ref(package) we provide a basic
description of how to quantify and control the approximation error.
Further, the RNA-seq example in Section @ref(examples) provides an
illustration of this in practice.

\emph{Step 3}

The final step implements a NORTA-inspired, Gaussian copula approach to
produce the desired margins. Steps 1 and 2 determine the MVN Pearson
correlation values that will eventually match the target correlation.
Then all that is required is a fast MVN simulator, a standard normal
CDF, and well-defined quantile functions for marginals. The MVN is
transformed to a copula (distribution with standard uniform margins) by
applying the normal CDF \(\phi(\cdot)\). The quantile functions
\(F^{-1}\) are applied across the margins to return the desired random
vector \({\bf Y}\).

\section[bigsimr-pkg]{The \pkg{bigsimr} R package}\label{package}

This section describes a low-dimensional (2D) random vector simulation
workflow via the \texttt{bigsimr} \texttt{R} package
(\url{https://github.com/SchisslerGroup/r-bigsimr}). The package
\texttt{bigsimr} provides an interface to the native code written in
Julia, registered as the \texttt{Bigsimr} Julia package
(\url{https://github.com/adknudson/Bigsimr.jl}). In addition to the
native Julia \texttt{Bigsimr} package and \texttt{R} interface
\texttt{bigsimr}, we also provide a python interface \texttt{bigsimr}
(\url{https://github.com/SchisslerGroup/python-bigsimr/}) that
interfaces with the Julia \texttt{Bigsimr} package. The Julia package
provides a high-performance implementation of our proposed random vector
generation algorithm and associated functions (see Section
@ref(algorithms)).

The subsections below describe the basic use of \texttt{bigsimr} by
stepping through an example workflow using the data set
\texttt{airquality} that contains daily air quality measurements in New
York, May to September 1973 \citep{Chambers1983}. This workflow proceeds
from setting up the computing environment, to data wrangling,
estimation, simulation configuration, random vector generation, and,
finally, result visualization.

\hypertarget{bivariate-example-description}{%
\subsection{Bivariate example
description}\label{bivariate-example-description}}

We illustrate the use of \texttt{bigsimr} motivated by the New York air
quality data set (\texttt{airquality}) included in the R
\texttt{datasets} package. First, we load the \texttt{bigsimr} library
and a few other convenient data science packages, including the
syntactically-elegant \texttt{tidyverse} suite of \texttt{R} packages.
The code chunk below prepares the computing environment:

\begin{CodeChunk}
\begin{CodeInput}
R> library("tidyverse")
R> library("bigsimr")
R> # Activate multithreading in Julia
R> Sys.setenv(JULIA_NUM_THREADS = parallel::detectCores())
R> # Load the Bigsimr and Distributions Julia packages
R> bs <- bigsimr_setup()
R> dist <- distributions_setup()
\end{CodeInput}
\end{CodeChunk}

For simplicity and to provide a minimal working example, we consider
bivariate simulation of the two \texttt{airquality} variables
\texttt{Temperature}, in degrees Fahrenheit, and \texttt{Ozone} level,
in parts per billion.

\begin{CodeChunk}
\begin{CodeInput}
R> df <- airquality %>% select(Temp, Ozone) %>% drop_na()
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeOutput}
Rows: 116
Columns: 2
$ Temp  <int> 67, 72, 74, 62, 66, 65, 59, 61, 74, 69, 66, 68, 58, 64, 66, 57, ~
$ Ozone <int> 41, 36, 12, 18, 28, 23, 19, 8, 7, 16, 11, 14, 18, 14, 34, 6, 30,~
\end{CodeOutput}
\end{CodeChunk}

Figure @ref(fig:ch030-aq-joint-dist) visualizes the bivariate
relationship between Ozone and Temperature. We aim to simulate random
two-component vectors mimicking this structure. The margins are not
normally distributed; particularly the Ozone level exhibits a strong
positive skew.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{article_bigsimr_files/figure-latex/ch030-aq-joint-dist-1}

}

\caption[Bivariate scatterplot of Ozone vs]{Bivariate scatterplot of Ozone vs. Temp with estimated marginal densities. We model the Ozone data as marginally log-normal and the Temperature data as normal.}\label{fig:ch030-aq-joint-dist}
\end{figure}
\end{CodeChunk}

Next, we specify the marginal distributions and correlation coefficient
(both type and magnitude). Here the analyst is free to be creative. For
this example, we avoid goodness-of-fit considerations to determine the
marginal distributions. But it seems sensible without domain knowledge
to estimate these quantities from the data, and \texttt{bigsimr}
contains fast functions designed for this task.

\hypertarget{specifying-marginal-distributions}{%
\subsection{Specifying marginal
distributions}\label{specifying-marginal-distributions}}

Based on the estimated densities in Figure
@ref(fig:ch030-aq-joint-dist), we assume \texttt{Temp} is normally
distributed and \texttt{Ozone} is log-normally distributed, as the
latter values are positive and skewed. We use the well-known, unbiased
estimators for the normal distribution's parameters and maximum
likelihood estimators for the log-normal parameters:

\begin{CodeChunk}
\begin{CodeInput}
R> df %>% select(Temp) %>% 
+   summarise_all(.funs = c(mean = mean, sd = sd))
\end{CodeInput}
\begin{CodeOutput}
   mean    sd
1 77.87 9.485
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> mle_mean <- function(x) mean(log(x))
R> mle_sd <- function(x) mean( sqrt( (log(x) - mean(log(x)))^2 ) )
R> df %>% 
+   select(Ozone) %>% 
+   summarise_all(.funs = c(meanlog = mle_mean, sdlog = mle_sd))
\end{CodeInput}
\begin{CodeOutput}
  meanlog  sdlog
1   3.419 0.6967
\end{CodeOutput}
\end{CodeChunk}

Next, we configure the input marginals for later input into
\texttt{rvec}. The marginal distributions are specified using
\texttt{Julia}'s \texttt{Distributions} package and stored in a vector.

\begin{CodeChunk}
\begin{CodeInput}
R> margins <- c(dist$Normal(mean(df$Temp), sd(df$Temp)),
+              dist$LogNormal(mle_mean(df$Ozone), mle_sd(df$Ozone)))
\end{CodeInput}
\end{CodeChunk}

\hypertarget{specifying-correlation}{%
\subsection{Specifying correlation}\label{specifying-correlation}}

As mentioned, the user must decide how to describe correlation based on
the particulars of the problem. For non-normal data and for improved
simulation accuracy/scalability in our scheme, we advocate the use of
Spearman's \(\rho\) correlation matrix \(R_S\) and Kendall's \(\tau\)
correlation matrix \(R_K\). We also support Pearson correlation
coefficient matching, while cautioning the user to check the performance
for their parametric multivariate model (see
\protect\hyperlink{simulations}{Monte Carlo evaluations} below for
evaluation strategies and guidance). These estimation methods are
classical approaches, not designed for high-dimensional correlation
estimation (see the \href{\%7B\#discussion}{Conclusion and Discussion}
sections for more on this).

\begin{CodeChunk}
\begin{CodeInput}
R> (R_S <- bs$cor(as.matrix(df), bs$Spearman))
\end{CodeInput}
\begin{CodeOutput}
      [,1]  [,2]
[1,] 1.000 0.774
[2,] 0.774 1.000
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{checking-target-correlation-matrix-admissibility}{%
\subsection{Checking target correlation matrix
admissibility}\label{checking-target-correlation-matrix-admissibility}}

First we use \texttt{cor\_bounds} to ensure that the pairwise target
correlation values are valid prior to the mapping step which constructs
\(R_X\) for the MVN input into NORTA (see Section @ref(algorithms)).
\texttt{cor\_bounds} estimates the pairwise lower and upper correlation
bounds using the Generate, Sort, and Correlate algorithm of
\citet{DH2011}.

\begin{CodeChunk}
\begin{CodeInput}
R> bounds <- bs$cor_bounds(margins)
R> bounds$lower
\end{CodeInput}
\begin{CodeOutput}
        [,1]    [,2]
[1,]  1.0000 -0.8745
[2,] -0.8745  1.0000
\end{CodeOutput}
\begin{CodeInput}
R> bounds$upper
\end{CodeInput}
\begin{CodeOutput}
       [,1]   [,2]
[1,] 1.0000 0.8748
[2,] 0.8748 1.0000
\end{CodeOutput}
\end{CodeChunk}

Since our single estimated Spearman correlation is within the
theoretical bounds, the correlation is valid as input to \texttt{rvec}.
But even if the 2-dimensional bounds are satisfied for each pair of
margins, this does not guarantee the feasibility of a \(d-\)variate
distribution \citep{BF17}.

To provide higher dimensional feasibility/admissibility checking, we
begin by mapping using \texttt{cor\_convert} and check admissibility.

\begin{CodeChunk}
\begin{CodeInput}
R> # Step 1. Mapping
R> (R_X <- bs$cor_convert(R_S, bs$Spearman, bs$Pearson))
\end{CodeInput}
\begin{CodeOutput}
       [,1]   [,2]
[1,] 1.0000 0.7886
[2,] 0.7886 1.0000
\end{CodeOutput}
\begin{CodeInput}
R> # Step 2. Check admissibility
R> bs$iscorrelation(R_X)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

The bounds on the Pearson correlation coefficient \(R_P\) between these
margins are restricted as seen below in our MC estimated correlation
bounds:

\begin{CodeChunk}
\begin{CodeInput}
R> bs$cor_bounds(margins[1], margins[2], bs$Pearson, n_samples = 1e6)
\end{CodeInput}
\begin{CodeOutput}
Julia Object of type NamedTuple{(:lower, :upper),Tuple{Float64,Float64}}.
(lower = -0.8799819883897544, upper = 0.8798667127065275)
\end{CodeOutput}
\end{CodeChunk}

Our MC estimate of the bounds slightly overestimates the theoretic
bounds of \((-0.881, 0.881)\). An analytic derivation is presented as an
Appendix.

\hypertarget{simulating-random-vectors}{%
\subsection{Simulating random vectors}\label{simulating-random-vectors}}

Finally, we arrive at the main function of \texttt{bigsimr}:
\texttt{rvec}. We now simulate \(B=10,000\) random vectors from the
assumed joint distribution of Ozone levels and Temp.

\begin{CodeChunk}
\begin{CodeInput}
R> x <- bs$rvec(10000, R_X, margins)
R> df_sim <- as.data.frame(x)
R> colnames(df_sim) <- colnames(df)
\end{CodeInput}
\end{CodeChunk}

Figure @ref(fig:ch030-plot-sim) plots the 10,000 simulated points.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{article_bigsimr_files/figure-latex/ch030-plot-sim-1}

}

\caption[Contour plot and marginal densities for the simulated bivariate distribution of Air Quality Temperatures and Ozone levels]{Contour plot and marginal densities for the simulated bivariate distribution of Air Quality Temperatures and Ozone levels. The simulated points mimic the observed data with respect to both the marginal characteristics and bivariate association.}\label{fig:ch030-plot-sim}
\end{figure}
\end{CodeChunk}

\hypertarget{simulations}{%
\section{Monte Carlo evaluations}\label{simulations}}

Before applying our methodology to real data simulation, we conduct
several Monte Carlo studies to investigate method performance. Since
marginal parameter matching in our scheme is essentially a sequence of
univariate inverse probability transforms, the challenging aspects are
the accuracy of dependency matching and computational efficiency at high
dimensions. To evaluate our methods in those respects, we design the
following numerical experiments to first assess accuracy of matching
dependency parameters in bivariate simulations and then time the
procedure in increasingly large dimension \(d\).

\hypertarget{bivariate-experiments}{%
\subsection{Bivariate experiments}\label{bivariate-experiments}}

We select bivariate simulation configurations to ultimately simulate our
motivating discrete-valued RNA-seq example, and, so we proceed in
increasing complexity, leading to the model in our motivating
application in Section @ref(examples). We begin with empirically
evaluating the dependency matching across all three supported
correlations --- Pearson's, Spearman's, and Kendall's --- in identical,
bivariate marginal configurations. For each pair of identical margins,
we vary the target correlation across \(\Omega\), the set of possible
admissible values for correlation type, to evaluate the simulation's
ability to obtain the theoretic bounds. The simulations progress from
bivariate normal, to bivariate gamma (non-normal yet continuous), and
bivariate negative binomial (mimicking RNA-seq counts).

Table @ref(tab:sims) lists our identical-marginal, bivariate simulation
configurations. We increase the simulate replicates \(B\) to check that
our results converge to the target correlations and gauge statistical
efficiency. We select distributions beginning with a standard
multivariate normal (MVN) as we expect the performance to be exact (up
to MC error) for all correlation types. Then, we select a non-symmetric
continuous distribution: a standard (rate =1) two-component multivariate
gamma (MVG). Finally, we select distributions and marginal parameter
values that are motivated by our RNA-seq data, namely values proximal to
probabilities and sizes estimated from the data (see
\href{examples}{Example applications} for estimation details). Thus we
arrive at a multivariate negative binomial (MVNB)
\(p_1 = p_2 = 3\times10^{-4}, r_1 = r_2 = 4, \rho \in \Omega\).

Table: (\#tab:sims) Identical margin, bivariate simulation
configurations to evaluate correlation matching accuracy and efficiency.

\begin{table}[]
\begin{tabular}{@{}lcr@{}}
\toprule
Simulation Reps $B$ & Correlation Types & Identical-margin 2D distribution \\ \midrule
$1000$ & Pearson ($\rho_P$) & ${\bf Y} \sim MVN( \mu= 0 , \sigma = 1, \rho_i ), i=1,\ldots,100$ \\
$10,000$ & Spearman ($\rho_S$) & ${\bf Y} \sim MVG( shape = 10, rate = 1, \rho_i ), i=1,\ldots,100$ \\
$100,000$ & Kendall ($\tau$) & ${\bf Y} \sim MVNB(p = 3\times10^{-4}, r = 4,\rho_i), i=1,\ldots,100$ \\ \bottomrule
\end{tabular}
\end{table}

For each of the unique 9 simulation configurations described above, we
estimate the correlation bounds and vary the correlations along a
sequence of 100 points evenly placed within the bounds, aiming to
explore \(\Omega\). Specifically, we set correlations
\(\{ \rho_1 = ( \hat{l} + \epsilon), \rho_2 = (\hat{l} + \epsilon) + \delta, \ldots, \rho_{100} = (\hat{u} - \epsilon) \}\),
with \(\hat{l}\) and \(\hat{u}\) being the estimated lower and upper
bounds, respectively, and increment value \(\delta\). The adjustment
factor, \(\epsilon=0.01\), is introduced to handle numeric issues when
the bound is specified exactly.

Figure @ref(fig:ch040-bPlot) displays the aggregated bivariate
simulation results. Table @ref(tab:ch040-BiError) contains the mean
absolute error (MAE) in reproducing the desired dependency measures for
the three bivariate scenarios.

\begin{CodeChunk}
\begin{table}

\caption{\label{tab:ch040-BiError}Average abolute error in matching the target dependency across the entire range of possible correlations for each bivariate marginal.}
\centering
\begin{tabular}[t]{lllr}
\toprule
No. of random vectors & Correlation type & Distribution & Mean abs. error\\
\midrule
1000 & Pearson & norm & 0.0168\\
1000 & Pearson & gamma & 0.0159\\
1000 & Pearson & nbinom & 0.0188\\
\addlinespace
1000 & Spearman & norm & 0.0203\\
1000 & Spearman & gamma & 0.0180\\
1000 & Spearman & nbinom & 0.0165\\
\addlinespace
1000 & Kendall & norm & 0.0118\\
1000 & Kendall & gamma & 0.0099\\
1000 & Kendall & nbinom & 0.0089\\
\addlinespace
10000 & Pearson & norm & 0.0048\\
10000 & Pearson & gamma & 0.0055\\
10000 & Pearson & nbinom & 0.0066\\
\addlinespace
10000 & Spearman & norm & 0.0058\\
10000 & Spearman & gamma & 0.0054\\
10000 & Spearman & nbinom & 0.0060\\
\addlinespace
10000 & Kendall & norm & 0.0034\\
10000 & Kendall & gamma & 0.0031\\
10000 & Kendall & nbinom & 0.0037\\
\addlinespace
1e+05 & Pearson & norm & 0.0016\\
1e+05 & Pearson & gamma & 0.0016\\
1e+05 & Pearson & nbinom & 0.0032\\
\addlinespace
1e+05 & Spearman & norm & 0.0018\\
1e+05 & Spearman & gamma & 0.0019\\
1e+05 & Spearman & nbinom & 0.0013\\
\addlinespace
1e+05 & Kendall & norm & 0.0011\\
1e+05 & Kendall & gamma & 0.0010\\
1e+05 & Kendall & nbinom & 0.0010\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

Overall, the studies show that our methodology is generally accurate
across the entire range of possible correlation values all three
dependency measures, at least in these limited simulation settings for
the rank-based correlations. Our Pearson matching performs nearly as
well as Spearman or Kendall, except for a slight increase in error for
negative binomial case. This is due the particularly large counts
generated with our choice of parameters for \(p\) and \(r\).

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{article_bigsimr_files/figure-latex/ch040-bPlot-1}

}

\caption[Bivariate simulations match target correlations across the entire range of feasible correlations]{Bivariate simulations match target correlations across the entire range of feasible correlations. The horizontal axis plots the specified target correlations for each bivariate margin. Normal margins are plotted in dark dark grey, gamma in medium grey, and negative binomial in light grey. As the number of simulated vectors $B$ increases from left to right, the variation in estimated correlations (vertical axis) decreases. The dashed line indicates equality between the specified and estimated correlations.}\label{fig:ch040-bPlot}
\end{figure}
\end{CodeChunk}

\hypertarget{scale-up-to-high-dimensions}{%
\subsection{Scale up to High
Dimensions}\label{scale-up-to-high-dimensions}}

With information of our method's accuracy from a low-dimensional
perspective, we now turn to assessing whether \texttt{bigsimr} can scale
to larger dimensional problems with practical computation times.
Specifically, we ultimately generate \(B=1,000\) random vectors for
\(d=\{100, 250, 500, 1000, 2500, 5000, 10000\}\) for each correlation
type, \(\{Pearson, Spearman, Kendall\}\) while timing the algorithm's
major steps.

To mimic the workflow, we first produce a synthetic ``data set'' by
completing the following steps:

\setstretch{1.5}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Produce heterogeneous gamma marginals by randomly selecting the
  \(j^{th}\) gamma shape parameter from
  \(U_j \sim uniform(1,10), j=1,\ldots,d\) and the \(j^{th}\) rate
  parameter from \(V_j \sim exp(1/5), j=1,\ldots,d\), with the constant
  parameters determined arbitrarily.
\item
  Produce a random full-rank Pearson correlation matrix via
  \texttt{cor\_randPD} of size \(d \times d\).
\item
  Simulate a ``data set'' of \(1,000 \times d\) random vectors via
  \texttt{rvec} (without matching the Pearson exactly). \setstretch{2.0}
\end{enumerate}

With the synthetic data set in hand, we complete and time the following
4 steps:

\setstretch{1.5}

\begin{enumerate}
\def\labelenumi{\roman{enumi}.}
\tightlist
\item
  Estimate the correlation matrix from the ``data'' in the \emph{Compute
  Correlation} step.
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Map the correlations to initialize the algorithm (Pearson to Pearson,
  Spearman to Pearson, or Kendall to Pearson) in the \emph{Adjust
  Correlation} step.
\item
  Check whether the mapping produces a valid correlation matrix and, if
  not, find the nearest PD correlation matrix in the \emph{Check
  Admissibility} step.
\item
  Simulate \(1,000\) vectors via the Inverse Transform method in the
  \emph{Simulate Data} step. \setstretch{2.0}
\end{enumerate}

The experiments are conducted on a MacBook Pro carrying a 2.4 GHz 8-Core
Intel Core i9 processor, with all 16 threads employed during
computation. The results for \(d \leq 500\) are fast with all times
(except Kendall) under 3 seconds. Table @ref(tab:ch040-moderateDtab)
displays the total computation time of the algorithms steps 1 through 3,
including estimation step i.

\begin{CodeChunk}

Table: Total time to produce 10,000 random vectors with a random correlation matrix and hetereogeneous gamma margins.

|Dimension |Correlation type | Total Time (Seconds)|
|:---------|:----------------|--------------------:|
|100       |Pearson          |                0.080|
|100       |Spearman         |                0.036|
|100       |Kendall          |                0.116|
|250       |Pearson          |                0.286|
|250       |Spearman         |                0.371|
|250       |Kendall          |                0.366|
|500       |Pearson          |                2.932|
|500       |Spearman         |                0.176|
|500       |Kendall          |                2.183|

\end{CodeChunk}

The results with larger dimensional vectors show scalability to
ultra-high dimensions for all three correlation types, although the
total times do become much larger. Figure @ref(fig:ch040-largeDfig)
displays computation times for \(d=\{1000, 2500, 5000, 10000\}\). For
\(d\) equal to 1000 and 2500, the total time is under a couple of
minutes. At \(d\) of 5000 and 10,000, Pearson correlation matching in
the \emph{Adjust Correlation} step becomes costly. Interestingly,
Pearson is actually faster than Kendall for \(d=10,000\) due to
bottlenecks in \emph{Compute Correlation} and \emph{Check
Admissibility}. Uniformly, matching Spearman correlations is faster,
with total times under 5 minutes for \(d=10,000\), making Spearman
correlations the most computationally-friendly dependency type. With
this in mind, we scaled the simulation to \(d=20,000\) for the Spearman
type and obtained the 1,000 vectors in under an hour (data not shown).
In principle, this would enable the simulation of an entire
human-derived RNA-seq data set. We note that for a given target
correlation matrix and margins, steps i, 1, and 2 only need to be
computed once and the third step, \emph{Simulate Data}, is fast under
all schemes for all \(d\) under consideration.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{article_bigsimr_files/figure-latex/ch040-largeDfig-1}

}

\caption[Computation times as d increases]{Computation times as d increases.}\label{fig:ch040-largeDfig}
\end{figure}
\end{CodeChunk}

\emph{Limitations, conclusions, and recommendations}

In the bivariate studies, we chose arbitrary simulation parameters for
three distributions, moving from the Gaussian to the discrete and
non-normal, multivariate negative binomial. Under these conditions, the
simulated random vectors sample the desired bivariate distribution
across the entire range of pairwise correlations for the three
dependency measures. The simulation results could differ for other
choices of simulation settings. Specifying extreme correlations near the
boundary or Frechet bounds could result in poor simulation performance.
Fortunately, it is straightforward to evaluate simulation performance by
using strategies similar to those completed above. We expect our random
vector generation to perform well for the vast majority of
NORTA-feasible correlation matrices, but advise to check the performance
before making inferences/further analyses.

Somewhat surprising, Kendall estimation and nearest PD computation scale
poorly compared to Spearman and, even, approximate Pearson matching. In
our experience, Kendall computation times are sensitive to the number of
cores, benefiting from multi-core parallelization. This could mitigate
some of the current algorithmic/implementation shortcomings. Despite
this, Kendall matching is still feasible for most high-dimensional data
sets. Finally, we note that one could use our single-pass algorithm
\texttt{cor\_fastPD} to produce a `close' (not nearest PD) to scale to
even higher dimensions with greater loss of accuracy.

\hypertarget{examples}{%
\section{RNA-seq data applications}\label{examples}}

This section demonstrates how to simulate multivariate data using
\texttt{bigsimr}, aiming to replicate the structure of high-dimensional
dependent count data. In an illustration of our proposed methodology, we
seek to simulate RNA-sequencing data by producing simulated random
vectors mimicking the observed data and its generating process. Modeling
RNA-seq using multivariate probability distributions is natural as
inter-gene correlation is an inherent part of biological processes
\citep{Wang2009b}. And yet, many models do not account for this, leading
to major disruptions to the operating characteristics of statistical
estimation, testing, and prediction. See \citet{Efron2012} for a
detailed discussion with related methods and \citet{Wu2012b},
\citet{Schissler2018}; \citet{Schissler2019} for applied examples. The
following subsections apply \texttt{bigsimr}'s methods to real RNA-seq
data, including replicating an estimated parametric structure, MC
probability estimation, and MC evaluation of correlation estimation
efficiency.

\hypertarget{simulating-high-dimensional-rna-seq-data}{%
\subsection{Simulating High-Dimensional RNA-seq
data}\label{simulating-high-dimensional-rna-seq-data}}

We begin by estimating the structure of the TCGA BRCA RNA-seq data set
(see \href{background}{Background}). Ultimately, we will simulate
\(B=10,000\) random vectors \({\bf Y}=(Y_1, \ldots, Y_d)^\top\) with
\(d=1000\). We assume a multivariate negative binomial (MVNB) model as
RNA-seq counts are often over-dispersed and correlated. Since all \(d\)
selected genes exhibit over-dispersion (data not shown), we proceed to
estimate the NB parameters \((r_i, p_i), i=1,\ldots,d\), to determine
the target marginal PMFs \(f_i\). To complete specification of the
simulation algorithm inputs, we estimate the Spearman correlation matrix
\({ \bf R}_{S}\) to characterize dependency.

With this goal in mind, we first estimate the desired correlation matrix
using the fast implementation provided by \texttt{bigsimr}:

\begin{CodeChunk}
\begin{CodeInput}
R> # Estimate Spearman's correlation on the count data
R> R_S <- bs$cor(as.matrix(brca1000), bs$Spearman)
\end{CodeInput}
\end{CodeChunk}

Next, we estimate the marginal parameters. We use the method of moments
(MoM) to estimate the marginal parameters for the multivariate negative
binomial model. While, marginal distributions are from the same
probability family (NB), they are heterogeneous in terms of the
parameters probability and size \((p_i, n_i)\) for \(i,\ldots,d\). The
functions below support this estimation for later use in \texttt{rvec}.

\begin{CodeChunk}
\begin{CodeInput}
R> make_nbinom_margins <- function(sizes, probs) {
+   margins <- lapply(1:length(sizes), function(i) {
+     dist$NegativeBinomial(sizes[i], probs[i])
+   })
+   do.call(c, margins)
+ }
\end{CodeInput}
\end{CodeChunk}

We apply these estimators to all 1000 genes across the 878 patients:

\begin{CodeChunk}
\begin{CodeInput}
R> nbinom_fit <- apply(brca1000, 2, mom_nbinom)
R> sizes <- nbinom_fit["size",]
R> probs <- nbinom_fit["prob",]
R> nb_margins <- make_nbinom_margins(sizes, probs)
\end{CodeInput}
\end{CodeChunk}

Notably, the estimated marginal NB probabilities \(\{ \hat{p}_i \}\) are
small --- ranging in the interval
\([\ensuremath{7.5305\times 10^{-7}} , \ensuremath{5.6592\times 10^{-4}}]\).
This gives rise to highly variable counts and, typically, less
restriction on potential pairwise correlation pairs. Once the functions
are defined/executed to complete marginal estimation, we specify targets
and generate the desired random vectors using \texttt{rvec}. Now we
check admissibility of the specified correlation matrix.

\begin{CodeChunk}
\begin{CodeInput}
R> # 1. Mapping step first
R> R_X <- bs$cor_convert(R_S, bs$Spearman, bs$Pearson)
R> # 2a. Check admissibility
R> (is_valid_corr <- bs$iscorrelation(R_X))
\end{CodeInput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\begin{CodeInput}
R> # 2b. compute nearest correlation
R> if (!is_valid_corr) {
+   R_X_pd <- bs$cor_nearPD(R_X)
+   ## Quantify the error
+   targets      <- R_X[lower.tri(R_X, diag = FALSE)]
+   approximates <- R_X_pd[lower.tri(R_X_pd, diag = FALSE)]
+   R_X          <- R_X_pd
+ }
R> summary(abs(targets - approximates))
\end{CodeInput}
\begin{CodeOutput}
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
0.000000 0.000201 0.000432 0.000534 0.000753 0.006931 
\end{CodeOutput}
\end{CodeChunk}

While the exact \(d \times d\) Spearman correlation matrix is not
strictly admissible in our scheme (as seen by the non-positive definite
result above), the approximation is close with a maximum absolute error
of 0.0069 and average absolute error of
\ensuremath{5.3366\times 10^{-4}} across the
\ensuremath{4.995\times 10^{5}} correlations.

\begin{CodeChunk}
\begin{CodeInput}
R> sim_nbinom <- bs$rvec(10000, R_X, nb_margins) 
R> colnames(sim_nbinom) <- colnames(brca1000)
\end{CodeInput}
\end{CodeChunk}

Figure @ref(fig:ch050-simDataFig) displays the simulated counts and
pairwise relationships for our example genes from Table
@ref(tab:ch010-realDataTab). Simulated counts roughly mimic the observed
data but with a smoother appearance due to the assumed parametric form
and with less extreme points then the observed data in Figure
@ref(fig:ch010-realDataFig). Figure @ref(fig:ch050-figBRCA) displays the
aggregated results of our simulation by comparing the specified target
parameter (horizontal axes) with the corresponding quantities estimated
from the simulated data (vertical axes). The evaluation shows that the
simulated counts approximately match the target parameters and exhibit
the full range of estimated correlation from the data.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.85\linewidth]{article_bigsimr_files/figure-latex/ch050-simDataFig-1}

}

\caption[Simulated data for three selected high-expressing genes generally replicates the estimated data structure]{Simulated data for three selected high-expressing genes generally replicates the estimated data structure. The data do not exhibit outlying points, but do possess the desired Spearman correlations, central tendencies, and discrete values.}\label{fig:ch050-simDataFig}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{fig/ch050-figBRCA}

}

\caption[Simulated random vectors from a multivariate negative binomial replicate the estimated structure from an RNA-seq data set]{Simulated random vectors from a multivariate negative binomial replicate the estimated structure from an RNA-seq data set. The dashed red lines indicate equality between estimated parameters from simulated data (vertical axes) and the specified target parameters (horizontal axes).}\label{fig:ch050-figBRCA}
\end{figure}
\end{CodeChunk}

\emph{Limitations, conclusions, and recommendations}

The results show overall good simulation performance for our choice of
parameters settings. Our settings were motivated by modeling
high-expressing genes from the TCGA BRCA data set. In general, the
ability to match marginal and dependence parameters depends on the
particular joint probability model. We recommend to evaluate and tune
your simulation until you can be assured of the accuracy.

\hypertarget{simulation-based-joint-probability-calculations}{%
\subsection{Simulation-based joint probability
calculations}\label{simulation-based-joint-probability-calculations}}

Many statistical tasks require evaluation of a joint probability mass
(or density) function:

\[
P( {\bf Y} = {\bf y} ), \: y_i \in \chi_i.
\]

where \(\chi_i\) is the sample space for the \(i^{th}\) component of the
random vector \(\bf{Y}\). Compact representations with convenient
computational forms are rare for high-dimensional constructions,
especially with heterogeneous, correlated marginal distributions (or
margins of mixed data types). Given a large number of simulated vectors
as produced above, estimated probabilities are readily given by counting
the proportion of simulated vectors meeting the desired condition. In
our motivating application, one may ask what is the probability that all
genes expressed greater than a certain threshold value \({\bf y}_0\).
This can be estimated as

\[
\hat{P}( {\bf Y} \ge {\bf y_0 } ) = \frac{1}{B} \sum_{b=1}^B I( {\bf Y}^{ (b) } \ge {\bf y_0} ),
\]

where \({\bf Y}^{(b)}\) is the \(b^{th}\) simulated vector from a total
of \(B\) simulation replicates and \(I(\cdot)\) is the indicator
function. For example, we can estimate from our \(B=10,000\) simulated
vectors that the probability of all genes expressing (i.e.,
\({\bf y}_i \geq 1, \forall \; i\)) is \(0.5979\).

\begin{CodeChunk}
\begin{CodeInput}
R> d <- ncol(sim_nbinom)
R> B <- nrow(sim_nbinom)
R> y0 <- 1
R> pHat <- mean(apply(sim_nbinom, 1, function(Y) {
+   all(Y >= y0)
+ }))
R> pHat
\end{CodeInput}
\begin{CodeOutput}
[1] 0.5979
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{evaluation-of-correlation-estimation-efficiency}{%
\subsection{Evaluation of correlation estimation
efficiency}\label{evaluation-of-correlation-estimation-efficiency}}

MC methods are routinely used in many statistical inferential tasks
including estimation, hypothesis testing, error rates, and empirical
interval coverage rates. To conclude the example applications, we
demonstrate how \texttt{bigsimr} can be used to evaluate estimation
efficiency. In particular, we wish to assess the error in our
correlation estimation above. We used a conventional method, based on
classical statistical theory which was not designed for high-dimensional
data. Indeed, high-dimensional covariance estimation (and precision
matrices) is an active area of statistical science, .e.g.,
\citep{Won2013g, VanWieringen2016}.

In this small example, we simulate \(m=30\) data sets with the number of
simulated vectors matching the number of patients in the BRCA data set,
\(N=878\). Since our simulation is much faster for the Pearson
correlation type (see Figure @ref(fig:ch040-largeDfig)), we only convert
the Spearman correlation matrix once (and ensure it is positive
definite). At each iteration, we estimate the quadratic loss (residual
sum of squared errors) from the specified \({\bf R}_{S}\), producing a
distribution of loss values.

\begin{CodeChunk}
\begin{CodeInput}
R> # Simulate random vectors equal to the sample size
R> N <- nrow(example_brca)
R> # create m random vectors and estimate correlation
R> simRho <- replicate(n = m, expr = {
+   tmpSim <- bs$rvec(N , R_X, nb_margins)
+   bs$cor(tmpSim, bs$Spearman)
+ }, simplify = FALSE)
R> # Evaluate the residual sum of squared error
R> sapply(simRho, function(R) sum((R - R_S)^2))
\end{CodeInput}
\end{CodeChunk}

The \texttt{R} summary function supplies the mean-augmented five-number
summary of the quadratic loss distribution computed above.

\begin{CodeChunk}
\begin{CodeInput}
R> summary(frobenius_loss)
\end{CodeInput}
\begin{CodeOutput}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    691     829     890     925     986    1224 
\end{CodeOutput}
\end{CodeChunk}

This distribution could be compared to high-dimensional designed
covariance estimators for guidance on whether the additional complexity
and computation time are warranted.

\hypertarget{discussion}{%
\section{Conclusion and discussion}\label{discussion}}

We have introduced a general-purpose high-dimensional multivariate
simulation algorithm and provide a user-friendly, high-performance
\texttt{R} package {[}\texttt{bigsimr}{]}. The random vector generation
method is inspired by NORTA \citep{Cario1997} and Gaussian copula-based
approaches \citep[\citet{BF17}, \citet{Xia17}]{MB13}. The major
contributions of this work are methods and software for flexible,
scalable simulation of HD multivariate probability distributions with
broad potential data analytic applications for modern, big-data
statistical computing. For example, one could simulate high-resolution
time series data, such as those consistent with an auto-regressive
moving average model exhibiting a specified Spearman structure. Or our
methods could be used to simulate sparsely correlated data, as many HD
methods assume, via specifying a \emph{spiked correlation matrix}.

It is customary to compare new tools and algorithms directly to existing
competing methods and software. In this study, however, we only employ
our proposed methodology, as our previous work has shown that existing
\texttt{R} tools are simply not designed to meet our high-dimensional
goal (see \citet{Li2019gpu} for evaluations of the \texttt{R}
\texttt{copula} package and others). For the bivariate simulations,
existing packages such as
\href{https://github.com/superdesolator/NORTARA}{\texttt{nortaRA}} work
well to match Pearson correlations exactly.

There are limitations to the methodology and implementation. We could
only investigate selected multivariate distributions in our Monte Carlo
studies. There may be instances that the methods do not perform well.
Along those lines, we expect that correlation values close to the
boundary of the feasible region could result in algorithm failure.
Another issue is that for discrete distributions, we use continuous
approximations when the support set is large. This could limit the
scalability/accuracy for particular discrete/mixed multivariate
distributions. Our method would also benefit computationally from faster
Kendall estimation and more finely tuned nearest PD calculation for
non-admissible Kendall matrices.

\newpage

\hypertarget{misc}{%
\section*{Supplementary Materials}\label{misc}}
\addcontentsline{toc}{section}{Supplementary Materials}

We provide an open-source implementation of our methodology as the
\texttt{Bigsimr.jl} Julia package, hosted on github at
\url{https://github.com/adknudson/Bigsimr.jl}. Additionally we provide R
and Python interfaces to \texttt{Bigsimr}, respectively at
\url{https://github.com/SchisslerGroup/r-bigsimr} and
\url{https://github.com/SchisslerGroup/python-bigsimr}.
\texttt{Bigsimr.jl} is an ongoing project, and feature requests or
issues can be submitted to
\url{https://github.com/adknudson/Bigsimr.jl/issues}.

\begin{CodeChunk}

\begin{center}\includegraphics[width=0.05\linewidth]{images/hex-bigsimr} \end{center}

\end{CodeChunk}

\hypertarget{acknowledgments}{%
\section*{Acknowledgment(s)}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgment(s)}

The authors gratefully acknowledge Heather Knudson's graphic design for
the \texttt{bigsimr} R Package. The results published here are in whole
or part based upon data generated by the TCGA Research Network:
\url{https://www.cancer.gov/tcga}.

\hypertarget{coi}{%
\section*{Disclosure statement}\label{coi}}
\addcontentsline{toc}{section}{Disclosure statement}

The authors report no conflict of interest.

\hypertarget{funding}{%
\section*{Funding}\label{funding}}
\addcontentsline{toc}{section}{Funding}

Research reported in this publication was supported by grants from the
National Institutes of General Medical Sciences (5 U54 GM104944,
GM103440) from the National Institutes of Health.

\newpage

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

\noindent Consider a bivariate example where we have a correlated
\((Y_1, Y_2)^\top\) with \(Y_1\sim N(\mu_1, \sigma_1^2)\) (normal) and
\(Y_2\sim LN(\mu_2, \sigma_2^2)\) (lognormal). First, let us note that
when we get such a vector from the \texttt{bigsimr} package then in fact
it can be represented as

\begin{equation}
(Y_1, Y_2)^\top \stackrel{d}{=} \left(X_1, e^{X_2}\right)^\top,
(\#eq:kram1)
\end{equation}

where \((X_1, X_2)^\top \sim N_2(\boldsymbol \mu, \boldsymbol \Sigma)\),
which is bivariate normal with mean vector
\(\boldsymbol \mu = (\mu_1, \mu_2)^\top\) and variance-covariance matrix

\begin{equation}
\boldsymbol \Sigma = 
\begin{bmatrix}
\sigma_1^2 & \rho \sigma_1\sigma_2\\
\rho \sigma_1\sigma_2 & \sigma_2^2
\end{bmatrix}
(\#eq:kram2)
\end{equation}

To see this, consider the three steps of the NORTA construction:

\begin{enumerate}

\item Generate $(Z_1, Z_2)^\top \sim N_2(\boldsymbol 0, \boldsymbol R)$, where

\begin{equation}
\boldsymbol R = 
\left[
\begin{array}{cc}
1 & \rho \\
\rho & 1
\end{array}
\right].
(\#eq:kram3)
\end{equation}

\item Transform $(Z_1, Z_2)^\top$ to $(U_1, U_2)^\top$ viz. $U_i =\Phi(Z_i)$,  $i=1,2$, where $\Phi(\cdot)$ is the standard normal CDF.

\item Return $(Y_1, Y_2)^\top$, where $Y_i=F_i^{-1}(U_i)$, $i=1,2$, and $F_i(\cdot)$ is the CDF of $Y_i$ and $F_i^{-1}(\cdot)$ is its inverse (the quantile function).

\end{enumerate}

In this example, \(F_1\) is the CDF of \(N(\mu_1, \sigma_1^2)\) while
\(F_2\) is the CDF of \(LN(\mu_2, \sigma_2^2)\), so that the two
required quantile functions are

\begin{equation}
F_1^{-1}(u) = \mu_1+\sigma_1 \Phi^{-1}(u)\,\,\, \mbox{and} \,\,\, F_2^{-1}(u) = e^{\mu_2+\sigma_2 \Phi^{-1}(u)}, 
(\#eq:kram4)
\end{equation}

as can be seen by standard calculation. Thus, when we apply Step 3 of
the NORTA algorithm, we get

\begin{equation}
F_1^{-1}(U_1) = \mu_1+\sigma_1 \Phi^{-1}(\Phi(Z_1)) =  \mu_1+\sigma_1 Z_1 \,\,\, \mbox{and} \,\,\, F_2^{-1}(U_2) = e^{\mu_2+\sigma_2 \Phi^{-1}(\Phi(Z_2))} = e^{\mu_2+\sigma_2 Z_2}, 
(\#eq:kram5)
\end{equation}

where
\((X_1, X_2)^\top = (\mu_1+\sigma_1 Z_1, \mu_2+\sigma_2 Z_2)^\top \sim N_2(\boldsymbol \mu, \boldsymbol \Sigma)\).
Consequently, the vector \((Y_1, Y_2)^\top\) obtained in Step 3 of the
algorithm has the structure provided by (\ref{kram1}).

\vspace{0.1in}

\noindent Next, we provide the exact covariance structure of the random
vector \((Y_1, Y_2)^\top\) given by (\ref{kram1}), and relate the
correlation of \(Y_1\) and \(Y_2\) to \(\rho\), which is the correlation
of the normal variables \(Z_1\), \(Z_2\) (and also \(X_1\) and \(X_2\)).
A straightforward albeit somewhat tedious algebra produces the following
result.

\begin{lemma}
Let ${\bf Y} = (Y_1, Y_2)^\top$ admit the stochastic representation (\ref{kram1}), where ${\bf X} = (X_1, X_2)^\top \sim N_2(\boldsymbol \mu, \boldsymbol \Sigma)$ with $\boldsymbol \mu$ and $\boldsymbol \Sigma$ as above. Then the mean vector and the variance-covariance matrix of ${\bf Y}$ are given by

\begin{equation}
\boldsymbol \mu_{{\bf Y}} = \left(\mu_1, e^{\mu_2+\sigma_2^2/2}\right)^\top
(\#eq:kram6)
\end{equation}

and

\begin{equation}
\boldsymbol \Sigma = 
\begin{bmatrix}
\sigma_1^2 & \rho \sigma_1\sigma_2  e^{\mu_2+\sigma_2^2/2}  \\
\rho \sigma_1\sigma_2 e^{\mu_2+\sigma_2^2/2} & \left[ e^{\mu_2+\sigma_2^2/2}\right]^2 \left[e^{\sigma_2^2} -1 \right],
\end{bmatrix}
(\#eq:kram7)
\end{equation}

respectively. 
\end{lemma}

\begin{proof}
The values of the means and the variances are obtained immediately from normal marginal distribution of $Y_1$ and lognormal marginal distribution of $Y_2$. It remains to establish the covariance of $Y_1$ and $Y_2$,

\begin{equation}
\mbox{Cov}(Y_1, Y_2) = \mathbb E(Y_1 Y_2) - \mathbb E(Y_1)\mathbb E(Y_2), 
(\#eq:covy1y2)
\end{equation}

and in particular the first expectation on the right-hand-side above. By using the tower property of expectations, the latter expectation can be expressed as

\begin{equation}
\mathbb E(Y_1 Y_2)  =  \mathbb E\left(X_1 e^{X_2}\right)  =  \mathbb E \left\{ \mathbb E \left(X_1 e^{X_2} | X_1 \right) \right\} = \mathbb E \left\{ X_1 \mathbb E \left(e^{X_2} | X_1 \right) \right\}.
(\#eq:kre1)
\end{equation}

Further, by using the fact that the conditional distribution of $X_2$ given $X_1=x_1$ is normal with mean $\mathbb E(X_2|X_1=x_1) = \mu_2+\rho\sigma_2(x_1-\mu_1)/\sigma_1$ and variance $\sigma_2^2(1-\rho^2)$, one can relate the inner expectation on the far right in (\ref{kre1}) to the moment generating function $M(t)$ of this conditional distribution evaluated at $t=1$, leading to

\begin{equation}
E \left(e^{X_2} | X_1 \right) = e^{ \mu_2+\rho\sigma_2(x_1-\mu_1)/\sigma_1 + \sigma_2^2(1-\rho^2)/2},
(\#eq:kre2)
\end{equation}

so that

\begin{equation}
\mathbb E(Y_1 Y_2)  =  e^{ \mu_2+ \frac{1}{2} \sigma_2^2(1-\rho^2)} \mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\}.
(\#eq:kre3)
\end{equation}

Since $X_1$ is normal with mean $\mu_1$ and variance $\sigma_1^2$, the expectation in (\ref{kre3}) becomes

\begin{equation}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = \int_{-\infty}^\infty x_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (x_1-\mu_1)} \frac{1}{\sqrt{2\pi}\sigma_1} e^{-\frac{1}{2\sigma_1^2}(x_1-\mu_1)^2}.
(\#eq:kre4)
\end{equation}

Upon substituting $u=x_1-\mu_1$, followed by some algebra, we arrive at

\begin{equation}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = e^{\frac{1}{2}\sigma_2^2\rho^2} \int_{-\infty}^\infty (u+\mu_1)  \frac{1}{\sqrt{2\pi}\sigma_1} e^{-\frac{1}{2\sigma_1^2}(u-\rho\sigma_1\sigma_2)^2}.
(\#eq:kre5)
\end{equation}

Since the integral in (\ref{kre5}) is the expectation $\mathbb E(X+\mu_1)$, where $X$ is normal with mean $\rho\sigma_1\sigma_2$ and variance $\sigma_1^2$, we conclude that

\begin{equation}
\mathbb E \left\{ X_1 e^{ \rho\frac{\sigma_2}{\sigma_1} (X_1-\mu_1)} \right\} = e^{\frac{1}{2}\sigma_2^2\rho^2} (\rho\sigma_1\sigma_2 +\mu_1),
(\#eq:kre6)
\end{equation}

which, in view of (\ref{kre3}), leads to

\begin{equation}
\mathbb E (Y_1Y_2) = e^{\mu_2 + \frac{1}{2}\sigma_2^2} (\rho\sigma_1\sigma_2 +\mu_1).
(\#eq:kre7)
\end{equation}

Finally, (\ref{covy1y2}), along with the expressions for the means of $Y_1$ and $Y_2$, produce the covariance of $Y_1$ and $Y_2$:

\begin{equation}
\mbox{Cov}(Y_1, Y_2) = e^{\mu_2 + \frac{1}{2}\sigma_2^2} (\rho\sigma_1\sigma_2 +\mu_1) - \mu_1 e^{\mu_2 + \frac{1}{2}\sigma_2^2} = \rho\sigma_1\sigma_2 e^{\mu_2 + \frac{1}{2}\sigma_2^2}.
(\#eq:kre8)
\end{equation}

This concludes the proof
\end{proof}

Using the above result, we can directly relate the correlation
coefficient of \(Y_1\) and \(Y_2\) with that of \(X_1\) and \(X_2\),

\begin{equation}
\rho_{Y_1, Y_2} = \rho \frac{\sigma_2}{\sqrt{e^{\sigma_2^2} -1}},
(\#eq:kram8)
\end{equation}

where \(\rho\) is the correlation of \(X_1\) and \(X_2\). The above
result is useful when studying the possible range of correlation of
\(Y_1\) and \(Y_2\) in the above example. It can be shown that the
factor on the far right in (\ref{kram8}) is a monotonically decreasing
function of \(\sigma_2\) on \((0,\infty)\), with the limits of 1 and 0
at zero and infinity, respectively. Thus, in principle, the range of
correlation of \(Y_1\) and \(Y_2\) is the same as that of \(X_1\) and
\(X_2\), as the factor on the far right in (\ref{kram8}) can be made
arbitrarily close to 1. However, by changing this factor we may affect
the marginal distributions of \(Y_1\) and \(Y_2\). It can be shown that
if the marginal distributions of \(Y_1\) and \(Y_2\) are fixed, then the
relation (\ref{kram8}) becomes

\begin{equation}
\rho_{Y_1, Y_2} = \rho \sqrt{\frac{\log(1+c_2^2)}{c_2^2}},
(\#eq:kram9)
\end{equation}

where \(c_2=\sigma_{Y_2}/\mu_{Y_2}\) is the \emph{coefficient of
variation} (CV) of the variable \(Y_2\). Thus, the range of possible
correlations in this model is not affected by the distribution of
\(Y_1\), and is determined by the CV of \(Y_2\) as follows:

\begin{equation}
- \sqrt{\frac{\log(1+c_2^2)}{c_2^2}} \leq \rho_{Y_1, Y_2} \leq \sqrt{\frac{\log(1+c_2^2)}{c_2^2}}. 
(\#eq:kram10)
\end{equation}

Plugging in the estimates of these quantities from Section
@ref(package), we see that

\begin{equation}
\frac{\hat{\sigma}_2}{\sqrt{e^{\hat{\sigma}_2^2} -1}} = 0.881.
(\#eq:kram11)
\end{equation}

Thus, the possible range of correlation becomes \((-0.881, 0.881)\).
This approximately agrees with the MC results provided.

\setstretch{1.0}

\bibliography{bigsimr.bib,packages.bib,alex.bib}

\end{document}