### 2021-06-01 060-conclusion-discussion.Rmd

# Conclusion and discussion {#discussion}

We developed a general-purpose high-dimensional multivariate simulation algorithm and provide a user-friendly, high-performance `R` package `bigsimr`, Julia package `Bigsimr` and Python package `bigsimr`. The random vector generation method is inspired by NORTA [@Cario1997] and Gaussian copula-based approaches [@MB13, @BF17, @Xia17]. The major contributions of this work are methods and software for flexible, scalable simulation of HD multivariate probability distributions with broad data analytic applications for modern, big-data statistical computing. For example, one could simulate high-resolution time series data, such as those consistent with an auto-regressive moving average model exhibiting a specified Spearman structure. Or our methods could be used to simulate sparsely correlated data, as many HD methods assume, via specifying a *spiked correlation matrix*.

<!-- 
It is customary to compare new tools and algorithms directly to existing competing methods and software. In this study, however, we only employ our proposed methodology, as our previous work has shown that existing `R` tools are simply not designed to meet our high-dimensional goal (see @Li2019gpu for evaluations of the `R` `copula` package and others). For the bivariate simulations, existing packages such as [`nortaRA`](https://github.com/superdesolator/NORTARA) work well to match Pearson correlations exactly.
-->

There are limitations to the methodology and implementation. We could only investigate selected multivariate distributions in our Monte Carlo studies. There may be instances that the methods do not perform well. Along those lines, we expect that correlation values close to the boundary of the feasible region could result in algorithm failure. Another issue is that for discrete distributions, we use continuous approximations when the support set is large. This could limit the scalability/accuracy for particular discrete/mixed multivariate distributions. Our method would also benefit computationally from faster Kendall estimation and more finely tuned nearest PD calculation for non-admissible Kendall matrices.