### 2021-06-25 070-supplementary-materials.Rmd

# Supplementary Materials {.unnumbered #misc}

We provide an open-source implementation of our methodology as the `Bigsimr.jl` Julia package, hosted on github at https://github.com/SchisslerGroup/Bigsimr.jl. Additionally we provide R and Python interfaces to `Bigsimr`, respectively on https://github.com/SchisslerGroup/r-bigsimr and https://github.com/SchisslerGroup/python-bigsimr. A stable release of the R package `bigsimr` is available on CRAN. `Bigsimr.jl` is an ongoing project, and feature requests or issues can be submitted to https://github.com/adknudson/Bigsimr.jl/issues.

```{r, out.width='5%'}
knitr::include_graphics('images/hex-bigsimr.png')
```

# Acknowledgment(s) {.unnumbered #acknowledgments}

The authors gratefully acknowledge Heather Knudson's graphic design for the `bigsimr` R Package. The results published here are in whole or part based upon data generated by the TCGA Research Network: https://www.cancer.gov/tcga.

# Disclosure statement {.unnumbered #coi}

The authors report no conflict of interest.

# Funding {.unnumbered #funding}

Research reported in this publication was supported by grants from the National Institutes of General Medical Sciences (5 U54 GM104944, GM103440) from the National Institutes of Health.

<!-- 
Nomenclature/Notation
-->